```{r}
library(tidyverse)
library(scales)
library(csgjcr)
library(highcharter)
library(htmlwidgets)
library(reactable)
library(reactablefmtr)

source("R/utils.R")
source("R/page-data.R")

# htmlwidgets seeds its element IDs from this, so rerenders give identical HTML
set.seed(2019)
```

```{r}
state_name_text <- "the United States"

srs_state_new <- read_srs_state()
latest_year <- max(srs_state_new$year)

### time series use an 11-year window ending in the latest data year
first_year <- latest_year - 10
srs_by_cat_us <- violent_totals(srs_state_new, first_year)
srs_us <- violent_by_offense(srs_state_new, first_year)

### US solve rates leave out the states excluded from solve rates
srs_solve_pool <- solve_rate_pool(srs_state_new)
srs_by_cat_us_solve <- violent_totals(srs_solve_pool, first_year)
srs_us_solve <- violent_by_offense(srs_solve_pool, first_year)

### violent crime counts and rates by state, one column per year
### solve rates are multiplied by 100 for highcharts formatting
### excluded states get missing solve rates, so the maps gray them out
srs_state_to_map <- srs_state_new |>
  filter(crime_cat == "Violent", year %in% c(base_year, latest_year)) |>
  group_by(state_name, state_abbr, pop_total, year) |>
  summarize(
    n_reported = sum(incidents_reported, na.rm = TRUE),
    n_solved = sum(incidents_cleared, na.rm = TRUE),
    .groups = "drop"
  ) |>
  mutate(
    period = if_else(year == base_year, "base", "latest"),
    viol_crime_rate_100k = (n_reported / pop_total) * 100000,
    solved_rate = n_solved / n_reported * 100
  ) |>
  pivot_wider(
    id_cols = c(state_name, state_abbr),
    names_from = period,
    values_from = c(n_reported, n_solved, viol_crime_rate_100k, solved_rate)
  ) |>
  mutate(
    solved_rate_percent_point_change = solved_rate_latest - solved_rate_base,
    viol_crime_rate_percent_change = (viol_crime_rate_100k_latest - viol_crime_rate_100k_base) / viol_crime_rate_100k_base * 100,
    across(starts_with("solved_rate_"), \(x) if_else(state_name %in% states_to_exclude_from_solved_rate_viz, NA_real_, x))
  )

### map tooltips, one column per map, and links to each state's page
### changes show a sign, e.g., "+5%" or "-12%"
srs_state_to_map <- srs_state_to_map |>
  mutate(
    no_solve_data = paste0(
      state_name, " did not report enough data", "<br>",
      "on solved crime to the FBI to", "<br>",
      "make reliable estimates."
    ),
    tooltip_crime_change = paste0(
      "<b>", state_name, "</b><br>",
      "Change in violent crime rate: ",
      sprintf("%+3d%%", as.numeric(comma(viol_crime_rate_percent_change, 1))), "<br>",
      "Violent incidents per 100k (", latest_year, "): ", comma(viol_crime_rate_100k_latest, 1), "<br>",
      "Violent incidents per 100k (", base_year, "): ", comma(viol_crime_rate_100k_base, 1)
    ),
    tooltip_crime_rate = paste0(
      "<b>", state_name, " – ", latest_year, "</b><br>",
      "Violent incidents per 100k: ", comma(viol_crime_rate_100k_latest, 1), "<br>",
      "Violent incidents reported: ", comma(n_reported_latest, 1), "<br>"
    ),
    tooltip_solve_change = if_else(
      state_name %in% states_to_exclude_from_solved_rate_viz, no_solve_data,
      paste0(
        "<b>", state_name, "</b><br>",
        "Change in solve rate: ",
        sprintf("%+3d", as.numeric(comma(solved_rate_percent_point_change, 1))), "<br>",
        "Solve rate (", latest_year, "): ", comma(solved_rate_latest, 1), "%", "<br>",
        "Solve rate (", base_year, "): ", comma(solved_rate_base, 1), "%", "<br>"
      )
    ),
    tooltip_solve_rate = if_else(
      state_name %in% states_to_exclude_from_solved_rate_viz, no_solve_data,
      paste0(
        "<b>", state_name, " – ", latest_year, "</b><br>",
        "Solve rate: ", comma(solved_rate_latest, 1), "%", "<br>",
        "Violent incidents reported: ", comma(n_reported_latest, 1), "<br>",
        "Violent incidents solved: ", comma(n_solved_latest, 1)
      )
    ),
    url = paste0("https://projects.csgjusticecenter.org/tools-for-states-to-address-crime/50-state-crime-data/?state=", tolower(state_abbr)),
    no_solve_data = NULL
  )

### describe a set of states for text, e.g., "12 states and the District of Columbia"
describe_states <- function(abbr) {
  n <- sum(abbr != "DC")
  paste0(n, if_else(n == 1, " state", " states"), if_else("DC" %in% abbr, " and the District of Columbia", ""))
}

states_rate_down <- srs_state_to_map |> filter(viol_crime_rate_percent_change < 0) |> pull(state_abbr)
states_rate_up <- srs_state_to_map |> filter(viol_crime_rate_percent_change > 0) |> pull(state_abbr)

states_solved_down <- srs_state_to_map |> filter(solved_rate_percent_point_change < 0) |> pull(state_abbr)
states_solved_up <- srs_state_to_map |> filter(solved_rate_percent_point_change > 0) |> pull(state_abbr)

### national violent crime rate change and offenses lower or higher than the base year
us_rate_change <- srs_by_cat_us |>
  filter(year %in% c(base_year, latest_year)) |>
  summarize(change = pct_change_first_last(incidents_reported_rate_total, year)) |>
  pull(change)

offense_rate_change <- srs_us |>
  filter(year %in% c(base_year, latest_year)) |>
  summarize(change = incidents_reported_rate_total[year == latest_year] - incidents_reported_rate_total[year == base_year], .by = group)

offenses_lower <- offense_rate_change |> filter(change < 0) |> pull(group) |> tolower()
offenses_higher <- offense_rate_change |> filter(change > 0) |> pull(group) |> tolower()

### import and prep hex map geo
hex <- sf::read_sf("data/us_states_hexgrid.geojson") |>
  select(state_abb = iso3166_2)

hex_gj <- hex |>
  sf::st_transform(3857) |>
  geojsonsf::sf_geojson() |>
  jsonlite::fromJSON(simplifyVector = FALSE)
```

```{r }
### national police protection expenditures in dollars of the latest survey year
asslgf_clean_national <- read_police_spending("US", read_inflation_index())

### national violent crime counts for officer workload
violent_us_by_year <- srs_state_new |>
  filter(crime_cat == "Violent") |>
  summarize(n_violent_reported = sum(incidents_reported, na.rm = TRUE), .by = year)

### officers per violent crime use the LEE population for the national crime rate
ucr_employee_data_national_clean <- read_lee_officers("US") |>
  left_join(violent_us_by_year, by = "year") |>
  mutate(violent_rate = n_violent_reported / pop_total) |>
  add_officer_rates() |>
  arrange(year)
```


```{r }
### import shr homicide incidents and clearances by state, year, and group
shr_state <- read_rds(csg_sp_path("jr_data_library", "data", "analysis", "fbi", "shr", "fbi_shr_state.rds"))

shr_years <- paste0(shr_first_year, "–", max(shr_state$year))

### pooled national solve rates for the shr_panels charts
shr_us <- function(category) {
  function_shr_grouping_for_national_plot(shr_state, category, shr_first_year)
}
```


# Trends in Violent Crime and Accountability

## 50-state violent crime data

::: {.panel-tabset}

## `r base_year`–`r latest_year` change

```{r}
#| fig-height: 5.5

state_map_specs <- list(
  crime_change = list(
    value = "viol_crime_rate_percent_change", tooltip = "tooltip_crime_change",
    title = paste0("Change in violent crime, ", base_year, "–", latest_year),
    subtitle = "Percent change in crime reported to police per 100,000 residents",
    caption = "FBI UCR Program, state crime estimates",
    colors = change_colors, label_format = "{value}%", legend_width = 250,
    accessibility = "state: {point.state_name}, percent change: {point.value:.1f}",
    symmetric = TRUE
  ),
  crime_rate = list(
    value = "viol_crime_rate_100k_latest", tooltip = "tooltip_crime_rate",
    title = paste0("Violent crime rate, ", latest_year),
    subtitle = "Crime reported to police per 100,000 residents",
    caption = "FBI UCR Program, state crime estimates",
    colors = rate_colors, label_format = "{value:,.0f}", legend_width = 250,
    accessibility = "state: {point.state_name}, violent crime rate: {point.value:.1f}"
  ),
  solve_change = list(
    value = "solved_rate_percent_point_change", tooltip = "tooltip_solve_change",
    title = paste0("Change in percentage of violent crime solved, ", base_year, "–", latest_year),
    subtitle = paste0("Percentage point change in rate of solved violent crimes from ", base_year, " to ", latest_year),
    caption = "FBI UCR Program, state crime estimates and clearances",
    colors = rev(change_colors), label_format = "{value}", legend_width = 300,
    accessibility = "state: {point.state_name}, percentage point change: {point.value:.1f}",
    symmetric = TRUE, null_interaction = TRUE
  ),
  solve_rate = list(
    value = "solved_rate_latest", tooltip = "tooltip_solve_rate",
    title = paste0("Solved violent crime, ", latest_year),
    subtitle = "Percentage of violent crimes solved",
    caption = "FBI UCR Program, state crime estimates and clearances",
    colors = rate_colors, label_format = "{value}%", legend_width = 400,
    accessibility = "state: {point.state_name}, solve rate: {point.value:.1f}%",
    null_interaction = TRUE
  )
)

state_map_chart(srs_state_to_map, hex_gj, state_map_specs$crime_change)
```

## `r latest_year` crime rate

```{r}
#| fig-height: 5.5

state_map_chart(srs_state_to_map, hex_gj, state_map_specs$crime_rate)
```

:::

The national violent crime rate was `r percent(abs(us_rate_change), 1, suffix = " percent")` `r if_else(us_rate_change < 0, "lower", "higher")` in `r latest_year` than in `r base_year`. In `r latest_year`, rates of `r knitr::combine_words(offenses_lower)` were lower than in `r base_year``r if (length(offenses_higher) > 0) paste0(", while the rate of ", knitr::combine_words(offenses_higher), " was higher")`.

Trends in violent crime also varied by state. Between `r base_year` and `r latest_year`, violent crime rates decreased in `r describe_states(states_rate_down)` and increased in `r describe_states(states_rate_up)`.

::: {.panel-tabset}

## `r base_year`–`r latest_year` change

```{r}
#| fig-height: 5.5

state_map_chart(srs_state_to_map, hex_gj, state_map_specs$solve_change)
```

## `r latest_year` solve rate

```{r}
#| fig-height: 5.5

state_map_chart(srs_state_to_map, hex_gj, state_map_specs$solve_rate)
```

:::

Between `r base_year` and `r latest_year`, the percentage of violent crimes solved by law enforcement increased in `r describe_states(states_solved_up)` and decreased in `r describe_states(states_solved_down)`.

`r knitr::combine_words(states_to_exclude_from_solved_rate_viz)` did not report enough data to the FBI on whether crimes were solved to make reliable estimates.

## Trends in reported violent crime

In `r max(srs_by_cat_us$year)`, `r srs_by_cat_us |> filter(year == max(year)) |> pull(incidents_reported_rate_total) |> comma(scale = 1e5)` violent crime incidents per 100,000 residents were reported to police in the United States.

::: {.panel-tabset}

## All violent crime

```{r}
us_violent_rates <- srs_by_cat_us |>
  mutate(
    incidents_reported_rate_total = incidents_reported_rate_total * 1e5,
    tooltip = paste0(
      "<b>", state_name, " – ", year, "</b><br>",
      "Violent incidents per 100k: ", comma(incidents_reported_rate_total, 1), "<br>",
      "Violent incidents reported: ", comma(incidents_reported, 1)
    )
  )

trend_chart(
  us_violent_rates, "incidents_reported_rate_total",
  paste0("Violent crime in ", state_name_text),
  "Crime reported to police per 100,000 residents",
  "FBI UCR Program, state crime estimates",
  value_label = "violent incidents per 100,000 residents",
  expansion = axis_expansion_mult, floor = NULL
)
```


```{r}
us_offense_rates <- srs_us |>
  left_join(offense_pal, by = c("group" = "crime")) |>
  group_by(year) |>
  mutate(tot_crime = sum(incidents_reported)) |>
  ungroup() |>
  mutate(
    pct = incidents_reported / tot_crime,
    incidents_reported_rate_total = incidents_reported_rate_total * 1e5,
    group = fct_reorder(group, incidents_reported_rate_total),
    tooltip = paste0(
      "<b>", group, " – ", year, "</b></br>",
      "Incidents reported per 100k: ", comma(incidents_reported_rate_total, 1), "<br>",
      "Incidents reported: ", comma(incidents_reported, 1), "<br>",
      "Total violent incidents reported: ", comma(tot_crime, 1), "<br>",
      "Share of all violent incidents: ", percent(pct, 1)
    )
  ) |>
  arrange(year, group)
```

## Homicide

```{r}
### homicide tooltips show one decimal because the rate is small
us_homicide_rates <- us_offense_rates |>
  filter(group == "Homicide") |>
  mutate(tooltip = paste0(
    "<b>", group, " – ", year, "</b></br>",
    "Incidents reported per 100k: ", comma(incidents_reported_rate_total, 0.1), "<br>",
    "Incidents reported: ", comma(incidents_reported, 1), "<br>",
    "Total violent incidents reported: ", comma(tot_crime, 1), "<br>",
    "Share of all violent incidents: ", percent(pct, 1)
  ))

offense_trend_chart(
  us_homicide_rates, "Homicide", paste0("Homicide in ", state_name_text),
  "Homicide incidents reported to police per 100,000 residents",
  offense_color("Homicide")
)
```

## Aggravated assault

```{r}
offense_trend_chart(
  us_offense_rates, "Aggravated assault", paste0("Aggravated assault in ", state_name_text),
  "Incidents reported to police per 100,000 residents where aggravated assault was the most serious offense",
  offense_color("Aggravated assault")
)
```

## Robbery

```{r}
offense_trend_chart(
  us_offense_rates, "Robbery", paste0("Robbery in ", state_name_text),
  "Incidents reported to police per 100,000 residents where robbery was the most serious offense",
  offense_color("Robbery")
)
```

## Rape

```{r}
offense_trend_chart(
  us_offense_rates, "Rape", paste0("Rape in ", state_name_text),
  "Incidents reported to police per 100,000 residents where rape was the most serious offense",
  offense_color("Rape")
)
```

## By offense

```{r}
us_offense_rates |>
  hchart(
    "column",
    stacking = "normal",
    groupPadding = 0.05,
    hcaes(year, incidents_reported_rate_total, group = group),
    accessibility = list(point = list(
      valueDescriptionFormat = "{point.series.name}, year: {point.x:.0f}, incidents per 100,000 residents: {point.y:.1f}"
    ))
  ) |>
  hc_colors(unique(us_offense_rates$color)) |>
  hc_title(text = paste0("Violent crime in ", state_name_text, " by offense")) |>
  hc_subtitle(text = "Crime reported to police per 100,000 residents") |>
  hc_caption(text = "FBI UCR Program, state crime estimates") |>
  hc_legend(reversed = TRUE) |>
  hc_setup()
```

:::


## Trends in solved violent crime

In `r max(srs_by_cat_us_solve$year)`, `r srs_by_cat_us_solve |> filter(year == max(year)) |> pull(pct_solved) |> percent(suffix = " percent")` of violent crime incidents reported to police in the United States were solved. The percentage of violent crimes that law enforcement solved fell from 2019 to 2022 and rose each year from 2023 to `r max(srs_by_cat_us_solve$year)`, reaching its highest level since at least `r min(srs_by_cat_us_solve$year)`.

::: {.panel-tabset}

## All violent crime

```{r}
us_violent_solve_rates <- srs_by_cat_us_solve |>
  filter(year != solve_rate_skip_year) |>
  mutate(
    pct_solved = pct_solved * 100,
    tooltip = paste0(
      "<b>", state_name, " – ", year, "</b><br>",
      "Solve rate: ", percent(pct_solved / 100, 1), "<br>",
      "Violent incidents reported: ", comma(incidents_reported, 1), "<br>",
      "Violent incidents solved: ", comma(incidents_solved, 1)
    )
  )

trend_chart(
  us_violent_solve_rates, "pct_solved",
  paste0("Solve rate of violent crime in ", state_name_text),
  caption = "FBI UCR Program, state crime estimates and clearances",
  value_label = "solve rate", value_format = "{point.y:.1f}%",
  expansion = axis_expansion_mult, ceiling = 100, y_format = "{value:,.0f}%"
)
```

## By offense

```{r}
us_offense_solve_rates <- srs_us_solve |>
  filter(year != solve_rate_skip_year) |>
  left_join(offense_pal, by = c("group" = "crime")) |>
  mutate(
    pct_solved = pct_solved * 100,
    group = fct_reorder(group, pct_solved, last),
    tooltip = paste0(
      "<b>", group, " – ", year, "</b></br>",
      "Solve rate: ", percent(pct_solved / 100, 1), "<br>",
      "Violent incidents reported: ", comma(incidents_reported, 1), "<br>",
      "Violent incidents solved: ", comma(incidents_solved, 1)
    )
  ) |>
  arrange(desc(year), pct_solved)

us_offense_solve_rates |>
  hchart(
    "spline",
    hcaes(year, pct_solved, group = group),
    accessibility = list(point = list(
      valueDescriptionFormat = "{point.series.name}, year: {point.x:.0f}, solve rate: {point.y:.1f}%"
    ))
  ) |>
  hc_colors(unique(us_offense_solve_rates$color)) |>
  hc_title(text = paste0("Solve rate of violent crime in ", state_name_text, " by offense")) |>
  hc_caption(text = "FBI UCR Program, state crime estimates and clearances") |>
  hc_setup() |>
  hc_yAxis(
    min = 0,
    max = 100,
    endOnTick = FALSE,
    labels = list(format = "{value:,.0f}%")
  )
```

:::

## Violent crime and solved violent crime by state, `r latest_year`

The following tables include violent crime rates and solve rates by state and offense. For more information on violent crime solve rates for each offense by state, [visit state-specific pages here](https://projects.csgjusticecenter.org/tools-for-states-to-address-crime/50-state-crime-data/).  

::: {.panel-tabset}

### Overall rates

```{r}
srs_state_to_map |>
  select(
    state_name,
    viol_crime_rate_100k_latest,
    n_reported_latest,
    viol_crime_rate_percent_change,
    solved_rate_latest,
    solved_rate_percent_point_change
  ) |>
  reactable_template(
    sort_col = list("state_name" = "asc"),
    columns = list(
      state_name = colDef(name = "State"),
      viol_crime_rate_100k_latest = colDef(name = paste0("Violent Incidents per 100k (", latest_year, ")")), 
      n_reported_latest = colDef(name = paste0("Violent Incidents Reported (", latest_year, ")")), 
      viol_crime_rate_percent_change = colDef(
        name = paste0("Percent change in violent crime rate (", base_year, " to ", latest_year, ")"),
        cell = add_plus_sign_percent_change
        ),
      solved_rate_latest = colDef(
        name = paste0("Solve Rate (", latest_year, ")"),
        format = colFormat(suffix = "%", digits = 0)
        ),
      solved_rate_percent_point_change = colDef(
        name = paste0("Percentage point change in solve rate (", base_year, " to ", latest_year, ")"),
        cell = add_plus_sign_percent_point_change,
        )
      )
    )

```



### By offense


`r solve_rate_legend("below", "indicates a solve rate less than the U.S. rate.")`

`r solve_rate_legend("above", "indicates a solve rate greater than or equal to the U.S. rate.")`


```{r}
### reported rates per 100k and solve rates by state and violent offense type
### solve rates can exceed 1 when a year clears older cases; excluded states
### and states with no reported incidents show blank solve rates
srs_state_by_offense_latest <- srs_state_new |>
  filter(crime_cat == "Violent", year == latest_year) |>
  mutate(
    solved_rate = (incidents_reported - incidents_unsolved) / incidents_reported,
    n_reported_rate = incidents_reported / pop_total * 100000
  ) |>
  pivot_wider(
    id_cols = state_name,
    names_from = group,
    values_from = c(n_reported_rate, solved_rate)
  ) |>
  janitor::clean_names() |>
  mutate(across(
    starts_with("solved_rate_"),
    \(x) if_else(state_name %in% states_to_exclude_from_solved_rate_viz | is.nan(x), NA_real_, x)
  ))

### each state is compared with the pooled US solve rate for the offense,
### leaving out the states excluded from solve rates
us_solved_rate_latest <- us_solve_benchmark(srs_state_new, latest_year)

solve_rate_layout <- list(
  aggravated_assault = list(offense = "Aggravated Assault"),
  homicide = list(offense = "Homicide", align = "center"),
  rape = list(offense = "Rape", min_width = 155, align = "left"),
  robbery = list(offense = "Robbery")
)

srs_state_by_offense_latest |>
  reactable_template(
    sort_col = list("state_name" = "asc"),
    columns = c(
      list(
        ### a right border sets off the sticky column
        state_name = colDef(
          name = "State", minWidth = 250, sticky = "left",
          style = list(borderRight = "1px solid #eee"),
          headerStyle = list(borderRight = "1px solid #eee")
        ),
        n_reported_rate_aggravated_assault = colDef(
          name = "Aggravated Assault Incidents per 100k",
          format = colFormat(digits = 0, separators = TRUE), minWidth = 120
        ),
        n_reported_rate_homicide = colDef(
          name = "Homicide Incidents per 100k",
          format = colFormat(digits = 0, separators = TRUE), align = "center"
        ),
        n_reported_rate_rape = colDef(
          name = "Rape Incidents per 100k",
          format = colFormat(digits = 0, separators = TRUE)
        ),
        n_reported_rate_robbery = colDef(
          name = "Robbery Incidents per 100k",
          format = colFormat(digits = 0, separators = TRUE)
        )
      ),
      solve_rate_columns(us_solved_rate_latest, "U.S. Rate", solve_rate_layout)
    )
  )
```

:::

Data regarding crime and solve rates come from the FBI's [Uniform Crime Reporting Program](https://www.fbi.gov/how-we-can-help-you/need-an-fbi-service-or-more-information/ucr). Violent crime includes homicide, aggravated assault, robbery, and rape. Crime rates are calculated by dividing the number of estimated incidents by the total population for a given geography. Solve rates are the number of reported crimes cleared by arrest or by exceptional means divided by the number of reported crimes. For more information, see the data sources and methodology section below. 

<br>

## Trends in law enforcement resources

State and local expenditures include spending on police, sheriffs, and highway patrols. The charts below also include staffing per 10,000 residents and per 1,000 violent crimes. These factors may show the level of policing a community receives, potentially influencing both crime and solve rates and providing insight into how law enforcement resources impact violent crime.

```{r}
### years and percent changes for the resource text
exp_min_year <- min(asslgf_clean_national$year)
exp_max_year <- max(asslgf_clean_national$year)

pct_change_spending <- asslgf_clean_national |>
  summarize(change = pct_change_first_last(n_inflation_adjusted_latest_year_dollars, year)) |>
  pull(change)

lee_min_year <- min(ucr_employee_data_national_clean$year)
lee_max_year <- max(ucr_employee_data_national_clean$year)

pct_change_staffing <- ucr_employee_data_national_clean |>
  summarize(change = pct_change_first_last(officer_pop_rate_per_10k, year)) |>
  pull(change)

pct_change_staffing_per_violent_offense <- ucr_employee_data_national_clean |>
  summarize(change = pct_change_first_last(officer_per_violent_offense_1k, year)) |>
  pull(change)
```

::: {.panel-tabset}

### State and local expenditures

Overall, state and local spending on law enforcement across the United States `r change_by(pct_change_spending)` from `r exp_min_year` to `r exp_max_year`, adjusting for inflation.

```{r}
spending_trend <- asslgf_clean_national |>
  mutate(
    tooltip = paste0(
      "<b>", "United States", " – ", year, "</b><br>",
      "Inflation-adjusted expenditure (", exp_max_year, " dollars): ",
      dollar(n_inflation_adjusted_latest_year_dollars, accuracy = 0.1, scale = 1e-9, suffix = "B")
    )
  )

trend_chart(
  spending_trend, "n_inflation_adjusted_latest_year_dollars",
  "State and local law enforcement expenditures",
  paste0("Law enforcement expenditures in ", exp_max_year, " dollars, ", exp_min_year, "–", exp_max_year),
  "U.S. Census Bureau, Annual Survey of State and Local Government Finances",
  value_label = paste0("inflation-adjusted expenditure in ", exp_max_year, " dollars"),
  value_format = "${point.y:,.0f}",
  y_formatter = "function() { return '$' + (this.value / 1000000000) + 'B'; }"
)
```

### Officers per capita

The number of officers per 10,000 residents in jurisdictions across the United States has `r change_by(pct_change_staffing)` from `r lee_min_year` to `r lee_max_year`.

```{r}
ucr_employee_data_national_clean |>
  mutate(tooltip = paste0(
    "<b>", "United States", " – ", year, "</b><br>",
    "Officers per 10k residents: ", comma(officer_pop_rate_per_10k, accuracy = .1)
  )) |>
  trend_chart(
    "officer_pop_rate_per_10k", "Officers per capita",
    paste0("Officers per 10k residents, ", lee_min_year, "–", lee_max_year),
    "FBI UCR Program, law enforcement employee counts and state crime estimates",
    value_label = "officers per 10,000 residents", expansion = axis_expansion_mult * 2.5
  )
```

### Officers per violent offense

The number of officers per 1,000 violent crimes in jurisdictions across the United States has `r change_by(pct_change_staffing_per_violent_offense)` from `r lee_min_year` to `r lee_max_year`.

```{r}
ucr_employee_data_national_clean |>
  mutate(tooltip = paste0(
    "<b>", "United States", " – ", year, "</b><br>",
    "Officers per 1k violent crimes: ", comma(officer_per_violent_offense_1k, accuracy = .1)
  )) |>
  trend_chart(
    "officer_per_violent_offense_1k", "Officers per violent offense",
    paste0("Officers per 1k violent crimes, ", lee_min_year, "–", lee_max_year),
    "FBI UCR Program, law enforcement employee counts and state crime estimates",
    value_label = "officers per 1,000 violent crimes", expansion = axis_expansion_mult
  )
```

:::

<br>

## Homicide solve rates by victim and incident characteristics

The following charts include homicide solve rates by race and ethnicity, gender, age, weapon type, and number of victims. For more information, see the data and methodology sections below.  

In the United States, homicide solve rates are highest when the victim is White and are lowest when the victim is Black. Homicide solve rates are highest when the victim is a woman and lowest when the victim is a man. Homicide solve rates are higher for victims 35 years and older.

::: {.panel-tabset}


### Race and ethnicity 

```{r}
shr_panel_chart(shr_panels$race, shr_us, shr_years)
```



### Gender

```{r}
shr_panel_chart(shr_panels$gender, shr_us, shr_years)
```



### Age


```{r}
shr_panel_chart(shr_panels$age, shr_us, shr_years)
```


### Weapon 

```{r}
shr_panel_chart(shr_panels$weapon, shr_us, shr_years)
```


### Number of victims

```{r}
shr_panel_chart(shr_panels$victims, shr_us, shr_years)
```

:::

<br>

```{r}
### states where agencies reporting at least one month cover less than
### low_coverage_threshold of the population, from R/01-prep-agency-county-data.R
state_coverage <- read_rds("data/state_coverage.rds")
coverage_year <- state_coverage$year[1]
low_coverage_states <- state_coverage |>
  filter(coverage < low_coverage_threshold) |>
  arrange(state_name) |>
  pull(state_name) |>
  as.character()

low_coverage_text <- if (length(low_coverage_states) > 0) {
  paste0(
    " Based on agency-level data from the Crime Data Explorer, agencies that reported at least one month in ",
    coverage_year, " covered less than ", percent(low_coverage_threshold, suffix = " percent"), " of the population in ",
    knitr::combine_words(low_coverage_states),
    ", so conclusions about crime changes in those states should be interpreted with additional caution."
  )
} else {
  ""
}
```

::: {.callout-note icon=false collapse=true}

## Data sources and methodology

Data regarding crime and solve rates come from the FBI's [Uniform Crime Reporting Program](https://www.fbi.gov/how-we-can-help-you/need-an-fbi-service-or-more-information/ucr){target="_parent"}. Counts of reported violent crime between `r min(srs_by_cat_us$year)` and `r latest_year` are derived from the Summary Reporting System (SRS) and the National Incident-Based Reporting System (NIBRS). The FBI estimates reported crime for each state from agency-level reports, accounting for agencies that did not report, and publishes these estimates through `r latest_year` on its [Crime Data Explorer](https://cde.ucr.cjis.gov/LATEST/webapp/#/pages/downloads){target="_parent"}. Solve rates apply each state's clearance rate among agencies that reported to the FBI, from the Crime Data Explorer, to the FBI's state estimates of reported crime.

The FBI stopped SRS data collection after data year 2020 and required law enforcement agencies to use NIBRS to report crime statistics to the FBI. This change decreased the number of agencies that submitted their crime data and thus reduced the reliability of state estimates in 2021, so the solve rate charts omit 2021. The FBI later allowed agencies unable to submit NIBRS data to submit SRS data again. In 2025, agencies reporting through NIBRS or SRS covered 96.2 percent of the U.S. population, and agencies reporting through NIBRS covered 89.5 percent, according to the FBI's [*Reported Crimes in the Nation, 2025*](https://cde.ucr.cjis.gov/LATEST/resources/reports/UCR_Summary_of_Reported_Crimes_in_the_Nation_2025.pdf){target="_parent"}.`r low_coverage_text`

Crime rates are calculated by dividing the number of estimated incidents by the total population. Population data come from the U.S. Census Bureau [Population Estimates Program](https://www.census.gov/programs-surveys/popest.html){target="_parent"}.

{{< include _methodology-exceptional-clearance.qmd >}}

U.S. solve rates from the FBI's state estimates leave out `r knitr::combine_words(states_to_exclude_from_solved_rate_viz)`, which did not report reliable clearance data, while U.S. crime rates include every state. U.S. homicide solve rates from the Supplementary Homicide Reports include every state.

Data regarding law enforcement expenditures were downloaded from the U.S. Census Bureau [Annual Survey of State and Local Government Finances](https://www.census.gov/programs-surveys/gov-finances.html){target="_parent"}. Expenditures include police protection spending by state governments and all local governments in each state, such as counties and municipalities. Expenditures are adjusted for inflation to constant dollars of the most recent year using the Chained Consumer Price Index for All Urban Consumers (C-CPI-U), following the [Census Bureau's guidance](https://www.census.gov/topics/income-poverty/income/guidance/current-vs-constant-dollars.html){target="_parent"}.

Officer employment analysis is restricted to sworn officers in state and local law enforcement agencies, as of October 31 of each year, and was downloaded from the FBI's [Crime Data Explorer law enforcement employee data](https://cde.ucr.cjis.gov/LATEST/webapp/#/pages/downloads){target="_parent"}. Officers per capita use the population served by agencies that reported employee counts. Officers per violent crime divide that rate by the national violent crime rate from the SRS state estimates.

{{< include _methodology-shr.qmd >}}

:::
