---
format:
  html:
    include-before-body: header.html
    css: styles.css
mainfont: "GT-America"
linkcolor: "#00355a"
execute:
  echo: false
  warning: false
  message: false
---

```{r}
library(tidyverse)
library(scales)
library(csgjcr)
library(highcharter)
library(htmlwidgets)
library(reactable)

source("R/utils.R")

# path to data folder on sharepoint
sp_path <- csg_sp_path("ad_hoc_requests/state_violent_crime_marshall/data")

state_name <- "United States"
state_name_text <- "the United States"
  
# read in crime data
viol_crime_us <- read_rds(file.path(sp_path, "viol_crime_us.rds"))
viol_crime_by_off_us <- read_rds(file.path(sp_path, "viol_crime_by_off_us.rds"))

viol_crime_by_state <- read_rds(file.path(sp_path, "viol_crime_by_state.rds"))

viol_crime_by_off_state <- read_rds(file.path(sp_path, "viol_crime_by_off_state.rds"))
viol_crime_2021_state <- read_rds(file.path(sp_path, "viol_crime_2021_state.rds"))
viol_crime_2021_by_off_state <- read_rds(file.path(sp_path, "viol_crime_2021_by_off_state.rds")) 

ucr_agency_pop <- read_rds(file.path(sp_path, "ucr_agency_pop.rds")) |> 
  filter(!state_abbr %in% c("GM", "PR", "VI"))

state_nibrs_pop_cov <- read_rds(file.path(sp_path, "state_nibrs_pop_cov.rds"))
# viol_prison_state <- read_rds("data/viol_prison_state.rds") |> 
#   filter(state == state_name)

# read in prison population data
us_prison <- read_rds(file.path(sp_path, "us_prison.rds"))
viol_prison_pop_state <- read_rds(file.path(sp_path, "viol_prison_pop_state.rds"))
viol_prison_pop_by_off_state <- read_rds(file.path(sp_path, "viol_prison_pop_by_off_state.rds"))

# check if we will consider this a nibrs state for this report
# which means 90% or more population coverage (from violent crime estimates)
nibrs_state <- viol_crime_2021_state |> 
  filter(pop_cov >= 0.9) |> 
  nrow() >= 1

# check if there is any ncrp data for this state
ncrp_state <- nrow(viol_prison_pop_state) > 1

# if we're going to use nibrs data join 2021 nibrs estimates with
# ucr srs state 2010-2020 - by offense
if (nibrs_state) {
  viol_crime_by_off_state <- viol_crime_by_off_state |> 
    bind_rows(viol_crime_2021_by_off_state)
  
    # and again just total violent crime
    viol_crime_by_state <- viol_crime_by_state |> 
    bind_rows(viol_crime_2021_state)
}

# create title of page
# we don't have IL clearance rate data so take that out of title
doc_title <- case_when(
  state_name == "Illinois" ~ paste(state_name, "data on violent crime and prison population"),
  ncrp_state ~ paste(state_name, "data on violent crime, arrests, and prison population"),
  TRUE ~ paste(state_name, "data on violent crime and arrests")
  )

# count how many agencies are nibrs reporters in the state
nibrs_agency_count <- ucr_agency_pop |> 
  count(is_nibrs) |> 
  arrange(is_nibrs)

# total number of agencies
total_agencies <- nrow(ucr_agency_pop)

# number of nibrs reporting agencies
yes_nibrs_agencies <- nibrs_agency_count$n[nibrs_agency_count$is_nibrs]

# number of non-nibrs agencies
no_nibrs_agencies <- total_agencies - yes_nibrs_agencies

nibrs_pop_cov_text <- percent(0.648, 1, suffix = " percent")

if (nibrs_pop_cov_text < total_agencies & nibrs_pop_cov_text == "100 percent") {
  nibrs_pop_cov_text <- "99 percent"
}

offense_pal <- tibble(
  color = jr_pal[1:4],
  crime = c("Homicide", "Robbery", "Rape", "Aggravated assault")
  )

prison_offense_pal <- tibble(
  color = jr_pal[1:5],
  off_comb = c("Homicide", "Robbery", "Rape", "Aggravated or simple assault",
               "Other violent offenses")
  )

```

---
title: "`r doc_title`"
---

## Violent crime in `r state_name_text`

```{r, results='asis'}
cat(paste0(
  "In 2021, ", comma(yes_nibrs_agencies, 1), " of ", comma(total_agencies, 1),
  " law enforcement agencies in ", state_name_text, " reported crime data to the FBI, covering ", nibrs_pop_cov_text, " of the nation's population. This was too few agencies to make reliable national esimates of violent crime in 2021, therefore the national crime data on this page end in 2020."
  ))
```

```{r, results='asis'}
cat(paste0(
  "A list of law enforcement agencies in ", state_name_text, " that did not fully report to the FBI on crime in 2021 can be found [below](#law-enforcement-agencies-not-reporting-to-nibrs)."
  ))
```

::: {.panel-tabset}

## All violent crime

```{r}
viol_crime_us |> 
  mutate(
    rate = actual_est / pop_total * 1e5,
    tooltip = paste0(
      "<b>", state, " – ", year, "</b><br>",
      "Violent offenses per 100k: ", comma(rate, 1), "<br>",
      "Violent offenses reported: ", comma(actual_est, 1)
      ),
    state = fct_relevel(state, "United States", after = Inf)
    ) |>
  hchart(
    "spline",
    hcaes(year, rate),
    color = jr_pal[1]
  ) |> 
  hc_title(text = paste("Violent crime in ", state_name_text, "reported to police per 100k residents")) |>
  hc_caption(text = paste0("FBI Uniform Crime Reporting Program SRS state estimates")) |> 
  hc_setup()
```

## By offense

```{r}
to_plot <- viol_crime_by_off_us |> 
  left_join(offense_pal, by = "crime") |> 
  group_by(year) |> 
  mutate(
    tot_crime = sum(actual_est),
    pct = actual_est / tot_crime,
    rate = actual_est / pop_total * 1e5
    ) |> 
  ungroup() |> 
  mutate(
    crime = fct_reorder(crime, rate),
    tooltip = paste0(
      "<b>", crime, " – ", year, "</b></br>",
      "Offenses reported per 100k: ", comma(rate, 1), "<br>",
      "Offenses reported: ", comma(actual_est, 1), "<br>",
      "Total violent offenses reported: ", comma(tot_crime, 1), "<br>",
      "Share of all violent offenses: ", percent(pct, 1)
      )
    ) |> 
  arrange(year, crime)

to_plot |> 
  hchart(
    "column",
    stacking = "normal",
    groupPadding = 0.05,
    hcaes(year, rate, group = crime),
  ) |> 
  hc_colors(unique(to_plot$color)) |> 
  hc_title(text = paste("Violent crime in ", state_name_text, " per 100k residents, by offense")) |>
  hc_caption(text = paste0("FBI Uniform Crime Reporting Program SRS state estimates")) |> 
  hc_setup()
```


## Homicide

```{r}
to_plot_2 <- to_plot |> 
  filter(crime == "Homicide")

to_plot_2 |> 
  hchart(
    "spline",
    stacking = "normal",
    hcaes(year, rate)
  ) |> 
  hc_yAxis(min = 0) |> 
  hc_colors(unique(to_plot_2$color)) |> 
  hc_title(text = paste("Homicide offenses in ", state_name_text, " per 100k residents")) |>
  hc_caption(text = paste0("FBI Uniform Crime Reporting Program SRS state estimates")) |> 
  hc_setup()
```

## Aggravated assault

```{r}
to_plot_3 <- to_plot |> 
  filter(crime == "Aggravated assault")

to_plot_3 |> 
  hchart(
    "spline",
    stacking = "normal",
    hcaes(year, rate)
  ) |> 
  hc_colors(unique(to_plot_3$color)) |> 
  hc_title(text = paste("Aggravated assault offenses in ", state_name_text, " per 100k residents")) |>
  hc_caption(text = paste0("FBI Uniform Crime Reporting Program SRS state estimates")) |> 
  hc_setup()
```

## Rape

```{r}
to_plot_4 <- to_plot |> 
  filter(crime == "Rape")

to_plot_4 |> 
  hchart(
    "spline",
    stacking = "normal",
    hcaes(year, rate)
  ) |> 
  hc_colors(unique(to_plot_4$color)) |> 
  hc_title(text = paste("Rape offenses in ", state_name_text, " per 100k residents")) |>
  hc_caption(text = paste0("FBI Uniform Crime Reporting Program SRS state estimates")) |> 
  hc_setup()
```

## Robbery

```{r}
to_plot_5 <- to_plot |> 
  filter(crime == "Robbery")

to_plot_5 |>
  hchart(
    "spline",
    stacking = "normal",
    hcaes(year, rate)
  ) |> 
  hc_colors(unique(to_plot_5$color)) |> 
  hc_title(text = paste("Robbery offenses in ", state_name_text, " per 100k residents")) |>
  hc_caption(text = paste0("FBI Uniform Crime Reporting Program SRS state estimates")) |> 
  hc_setup()
```

:::

#### Background

Beginning with data year 2021, the FBI discontinued collecting Uniform Crime Reporting data via the Summary Reporting System and required law enforcement agencies to use the [National Incident-Based Reporting System (NIBRS)](https://bjs.ojp.gov/national-incident-based-reporting-system-nibrs) to report crime statistics to the FBI. Because many states did not have full NIBRS reporting coverage in 2021, the FBI and Bureau of Justice Statistics [generated estimated crime statistics](https://bjs.ojp.gov/library/publications/estimation-procedures-crimes-united-states-based-nibrs-data) based on the agencies that reported data to the FBI. The reported crime and clearance rates for 2021 have a level of uncertainty represented by a margin of error for each estimate because not all agencies reported data to the FBI.

### Clearance rate of violent crime, 2020

```{r}
#| column: page-inset
#| out.width: "990px"
#| out.height: "600px"
hex <- sf::read_sf("us_states_hexgrid.geojson") |> 
  select(state_abb = iso3166_2)

hex_gj <- hex %>% 
  sf::st_transform(3857) %>%
  geojsonsf::sf_geojson() %>% 
  jsonlite::fromJSON(simplifyVector = FALSE)

highchart_hex_map <- function(x) {
  highchart()  |> 
    hc_add_series_map(
      map = hex_gj,
      df = x,
      joinBy = "state_abb",
      value = "rate",
      dataLabels = list(enabled = TRUE, format = "{point.state_abb}",
                        style = list(fontSize = "12px", fontWeight = "regular",
                                     textOutline = 0)),
      nullColor = "#e8e8e8"
      )  |> 
    hc_colorAxis(minColor = "#f7fdff", maxColor = "#1795BF") |> 
    hc_setup()
}

to_plot <- viol_crime_by_state |> 
  filter(year == 2020) |> 
  mutate(
    clearance_rate = if_else(state_abb == "IL", NA_real_, clearance_rate),
    clearance_rate = clearance_rate * 100,
    rate = clearance_rate,
    tooltip = paste0(
      "<b>", state, " – ", year, "</b><br>",
      "Clearance rate: ", percent(clearance_rate / 100, 1), "<br>",
      "Violent offenses reported: ", comma(actual_est, 1), "<br>",
      "Violent offenses cleared: ", comma(cleared_est, 1)
      ),
    url = paste0("state-viol-crime-", tolower(state_abb), ".html")
    )

highchart_hex_map(to_plot) |> 
  hc_tooltip(enabled = FALSE) |> 
  hc_plotOptions(
    series = list(
      point = list(
        events = list(
          click = JS("function(){location.href = this.options.url}"))
        )
      )
    )
```

## Law enforcement agencies not reporting to NIBRS

Not all law enforcement agencies have begun reporting NIBRS data to the FBI. As of data year 2021, `r comma(yes_nibrs_agencies, 1)` of `r comma(total_agencies, 1)` agencies in `r  state_name_text` reported NIBRS data. Agencies that *did not* report in 2021 are listed in the table below. The population column is the number of residents of the jurisdiction covered by the law enforcement agency.[^zero]

[^zero]: Some agencies, such as university and park police, are considered to have a population of zero by the FBI."


```{r}
reactable_template <- function(df, sort_col = "rate", ...) {
  reactable(
    df,
    highlight = TRUE,
    searchable = TRUE,
    defaultSorted = sort_col,
    showPageSizeOptions	= TRUE,
    pageSizeOptions = c(10, 25, 100),
    defaultColDef = colDef(
      vAlign = "center", 
      format = colFormat(digits = 0, separators = TRUE),
      headerStyle = list(fontWeight = 700, fontFamily = default_fonts,
                         fontVariant = "all-petite-caps"),
      style = list(fontWeight = 400, fontFamily = default_fonts)
      ),
    ...
      )
}

ucr_agency_pop |> 
  filter(!is_nibrs) |> 
  mutate(county_name = if_else(county_name == "Not Specified", "", county_name)) |> 
  select(agency_name, state_abbr, county_name, population) |> 
  arrange(desc(population)) |>
  reactable_template(
    sort_col = list("population" = "desc"),
    columns = list(
      agency_name = colDef(name = "Agency", minWidth = 250),
      state_abbr = colDef(name = "State"),
      county_name = colDef(name = "County"),
      population = colDef(name = "Population")
    )
  )
```
