---
format:
  html:
    include-before-body: header.html
    css: styles.css
mainfont: "GT-America"
linkcolor: "#00355a"
execute:
  echo: false
  warning: false
  message: false
---

```{r}
library(tidyverse)
library(scales)
library(csgjcr)
library(highcharter)
library(htmlwidgets)
library(reactable)

source("R/utils.R")

# path to data folder on sharepoint
sp_path <- csg_sp_path("ad_hoc_requests/state_violent_crime_marshall/data")

state_name <- "United States"
state_name_text <- "the United States"
  
# read in crime data
viol_crime_us <- read_rds(file.path(sp_path, "viol_crime_us.rds"))
viol_crime_by_off_us <- read_rds(file.path(sp_path, "viol_crime_by_off_us.rds"))

viol_crime_by_state <- read_rds(file.path(sp_path, "viol_crime_by_state.rds"))

viol_crime_by_off_state <- read_rds(file.path(sp_path, "viol_crime_by_off_state.rds"))
viol_crime_2021_state <- read_rds(file.path(sp_path, "viol_crime_2021_state.rds"))
viol_crime_2021_by_off_state <- read_rds(file.path(sp_path, "viol_crime_2021_by_off_state.rds")) 

ucr_agency_pop <- read_rds(file.path(sp_path, "ucr_agency_pop.rds")) |> 
  filter(!state_abbr %in% c("GM", "PR", "VI"))

state_nibrs_pop_cov <- read_rds(file.path(sp_path, "state_nibrs_pop_cov.rds"))
# viol_prison_state <- read_rds("data/viol_prison_state.rds") |> 
#   filter(state == state_name)

# read in prison population data
us_prison <- read_rds(file.path(sp_path, "us_prison.rds"))
viol_prison_pop_state <- read_rds(file.path(sp_path, "viol_prison_pop_state.rds"))
viol_prison_pop_by_off_state <- read_rds(file.path(sp_path, "viol_prison_pop_by_off_state.rds"))

# create title of page
# we don't have IL clearance rate data so take that out of title
doc_title <- paste(state_name, "data on violent crime and arrests")

# count how many agencies are nibrs reporters in the state
nibrs_agency_count <- ucr_agency_pop |> 
  count(is_nibrs) |> 
  arrange(is_nibrs)

# total number of agencies
total_agencies <- nrow(ucr_agency_pop)

# number of nibrs reporting agencies
yes_nibrs_agencies <- nibrs_agency_count$n[nibrs_agency_count$is_nibrs]

# number of non-nibrs agencies
no_nibrs_agencies <- total_agencies - yes_nibrs_agencies

nibrs_pop_cov_text <- percent(0.648, 1, suffix = " percent")


min_max_viol_crime <- viol_crime_us |> 
  filter(year %in% c(min(year), max(year))) |> 
  arrange(year)

max_year <- min_max_viol_crime$year[2]
min_year <- min_max_viol_crime$year[1] 

pct_change <- (min_max_viol_crime$actual_est[2] - min_max_viol_crime$actual_est[1]) / min_max_viol_crime$actual_est[1]

change_text <- case_when(
  pct_change > 0.005  ~ paste0(percent(abs(pct_change), 1),
                                    " more"),
  pct_change < -0.005 ~ paste0(percent(abs(pct_change), 1),
                                    " fewer"),
  TRUE                ~ "nearly the same number of"
  )

pct_change_clear <- min_max_viol_crime$clearance_rate[2] - min_max_viol_crime$clearance_rate[1]

change_text_clear <- case_when(
  pct_change_clear > 0.005 &
    pct_change_clear < 0.015 ~ paste0(comma(abs(pct_change_clear * 100), 1),
                                    " percentage point higher than"),
  pct_change_clear >= 0.015  ~ paste0(comma(abs(pct_change_clear * 100), 1),
                                    " percentage points higher than"),
  pct_change_clear < -0.005 &
    pct_change_clear > -0.015 ~ paste0(comma(abs(pct_change_clear * 100), 1),
                                    " percentage point lower than"),
  pct_change_clear <= -0.015 ~ paste0(comma(abs(pct_change_clear * 100), 1),
                                    " percentage points lower than"),
  TRUE                      ~ "nearly the same as"
  )

max_viol_crime_off <- viol_crime_by_off_us |> 
  filter(year == max(year)) |> 
  arrange(desc(actual_est)) |> 
  mutate(pct = actual_est / sum(actual_est))
```

---
title: "`r doc_title`"
---

## Violent crime in `r state_name_text`

```{r, results='asis'}
cat(paste0(
  "In 2021, ", comma(yes_nibrs_agencies, 1), " of ", comma(total_agencies, 1),
  " law enforcement agencies in ", state_name_text, " reported crime data to the FBI, covering ", nibrs_pop_cov_text, " of the nation's population. This was too few agencies to make reliable national esimates of violent crime in 2021, therefore the national crime data presented on this page end in 2020."
  ))
```

```{r, results='asis'}
cat(paste0(
  "A list of law enforcement agencies in ", state_name_text, " that did not fully report to the FBI on crime in 2021 can be found [below](#law-enforcement-agencies-not-reporting-to-nibrs)."
  ))
```

::: {.panel-tabset}

## All violent crime

```{r}
viol_crime_us |> 
  mutate(
    rate = actual_est / pop_total * 1e5,
    tooltip = paste0(
      "<b>", state, " – ", year, "</b><br>",
      "Violent offenses per 100k: ", comma(rate, 1), "<br>",
      "Violent offenses reported: ", comma(actual_est, 1)
      ),
    state = fct_relevel(state, "United States", after = Inf)
    ) |>
  hchart(
    "spline",
    hcaes(year, rate),
    color = jr_pal[1]
  ) |> 
  hc_title(text = paste0("Violent crime in ", state_name_text)) |>
  hc_subtitle(text = "Crime reported to police per 100,000 residents") |> 
  hc_caption(text = "FBI <em>Crime in the U.S.</em>") |> 
  hc_setup()
```


```{r}
to_plot <- viol_crime_by_off_us |> 
  left_join(offense_pal, by = "crime") |> 
  group_by(year) |> 
  mutate(
    tot_crime = sum(actual_est),
    pct = actual_est / tot_crime,
    rate = actual_est / pop_total * 1e5
    ) |> 
  ungroup() |> 
  mutate(
    crime = fct_reorder(crime, rate),
    tooltip = paste0(
      "<b>", crime, " – ", year, "</b></br>",
      "Offenses reported per 100k: ", comma(rate, 1), "<br>",
      "Offenses reported: ", comma(actual_est, 1), "<br>",
      "Total violent offenses reported: ", comma(tot_crime, 1), "<br>",
      "Share of all violent offenses: ", percent(pct, 1)
      )
    ) |> 
  arrange(year, crime)
```

## Homicide

```{r}
to_plot_2 <- to_plot |> 
  filter(crime == "Homicide")

to_plot_2 |> 
  hchart(
    "spline",
    stacking = "normal",
    hcaes(year, rate)
  ) |> 
  hc_yAxis(min = 0) |> 
  hc_colors(unique(to_plot_2$color)) |> 
  hc_title(text = paste0("Homicide in ", state_name_text)) |>
  hc_subtitle(text = "Homicide offenses reported to police per 100,000 residents") |> 
  hc_caption(text = "FBI <em>Crime in the U.S.</em>") |>
  hc_setup()
```

## Aggravated assault

```{r}
to_plot_3 <- to_plot |> 
  filter(crime == "Aggravated assault")

to_plot_3 |> 
  hchart(
    "spline",
    stacking = "normal",
    hcaes(year, rate)
  ) |> 
  hc_colors(unique(to_plot_3$color)) |> 
  hc_title(text = paste0("Aggravated assault in ", state_name_text)) |>
  hc_subtitle(text = "Aggravated assault offenses reported to police per 100,000 residents") |>
  hc_caption(text = "FBI <em>Crime in the U.S.</em>") |> 
  hc_setup()
```

## Robbery

```{r}
to_plot_5 <- to_plot |> 
  filter(crime == "Robbery")

to_plot_5 |>
  hchart(
    "spline",
    stacking = "normal",
    hcaes(year, rate)
  ) |> 
  hc_colors(unique(to_plot_5$color)) |> 
  hc_title(text = paste0("Robbery in ", state_name_text)) |>
  hc_subtitle(text = "Robbery offenses reported to police per 100,000 residents") |>
  hc_caption(text = "FBI <em>Crime in the U.S.</em>") |> 
  hc_setup()
```

## Rape

```{r}
to_plot_4 <- to_plot |> 
  filter(crime == "Rape")

to_plot_4 |> 
  hchart(
    "spline",
    stacking = "normal",
    hcaes(year, rate)
  ) |> 
  hc_colors(unique(to_plot_4$color)) |> 
  hc_title(text = paste0("Rape in ", state_name_text)) |>
  hc_subtitle(text = "Rape offenses reported to police per 100,000 residents") |>
  hc_caption(text = "FBI <em>Crime in the U.S.</em><br>The FBI <a href='https://ucr.fbi.gov/recent-program-updates/new-rape-definition-frequently-asked-questions'>broadened the definition of rape<a/> for crime reporting in 2013") |> 
  hc_setup()
```

## By offense

```{r}
to_plot |> 
  hchart(
    "column",
    stacking = "normal",
    groupPadding = 0.05,
    hcaes(year, rate, group = crime),
  ) |> 
  hc_colors(unique(to_plot$color)) |> 
  hc_title(text = paste0("Violent crime in ", state_name_text, ", by offense")) |>
  hc_subtitle(text = "Crime reported to police per 100,000 residents") |> 
  hc_caption(text = "FBI <em>Crime in the U.S.</em>") |> 
  hc_setup()
```

:::

::: {.callout-note icon=false collapse=true}

## Background on crime data reporting

Beginning with data year 2021, the FBI discontinued collecting Uniform Crime Reporting data via the Summary Reporting System and required law enforcement agencies to use the [National Incident-Based Reporting System (NIBRS)](https://bjs.ojp.gov/national-incident-based-reporting-system-nibrs) to report crime statistics to the FBI. Because many states did not have full NIBRS reporting coverage in 2021, the FBI and Bureau of Justice Statistics [generated estimated crime statistics](https://bjs.ojp.gov/library/publications/estimation-procedures-crimes-united-states-based-nibrs-data) based on the agencies that reported data to the FBI. The reported crime and clearance rates for 2021 have a level of uncertainty represented by a margin of error for each estimate because not all agencies reported data to the FBI.

:::

## Arrests for violent crime in `r state_name_text`


```{r}
clear_rate_max <- paste0(comma(min_max_viol_crime$clearance_rate[2] * 100, 1), " percent")
```

In `r max_year`, `r clear_rate_max` of violent crimes reported to police were cleared by arrest or exceptional means.[^exceptional] This is `r change_text_clear` the clearance rate in `r min_year`.

[^exceptional]: An incident is considered to be [cleared by exceptional means](https://ucr.fbi.gov/crime-in-the-u.s/2019/crime-in-the-u.s.-2019/topic-pages/clearances) if a law enforcement agency is unable to arrest or formally charge a person with a crime, but the agency has met the following conditions:
    
    1) Identified the person who committed the crime
    2) Gathered enough evidence to support an arrest, make a charge, and turn over the person to the court for prosecution
    3) Identified the suspect’s exact location so that the person could be taken into custody immediately
    4) Encountered a circumstance outside the control of law enforcement that prohibits the agency from arresting, charging, and prosecuting the person
    
    Examples of exceptional clearances include the death of the person who committed the crime; the victim’s refusal to cooperate with the prosecution after the person who committed the crime has been identified; or the denial of extradition.

::: {.panel-tabset}

## All violent crime

```{r}
viol_crime_us |> 
  mutate(
    clearance_rate = clearance_rate * 100,
    tooltip = paste0(
      "<b>", state, " – ", year, "</b><br>",
      "Clearance rate: ", percent(clearance_rate / 100, 1), "<br>",
      "Violent offenses reported: ", comma(actual_est, 1), "<br>",
      "Violent offenses cleared: ", comma(cleared_est, 1)
      ),
    state = fct_relevel(state, "United States", after = Inf)
    ) |> 
  hchart(
    "spline",
    hcaes(year, clearance_rate),
    color = jr_pal[1],
    dashStyle = "solid",
    opacity = c(1, 0.75)
  ) |> 
  hc_title(text = paste("Violent crime clearance rate in ", state_name_text)) |>
  hc_caption(text = "FBI <em>Crime in the U.S.</em>") |> 
  hc_setup() |> 
  hc_yAxis(labels = list(format = "{value:,.0f}%"))
```

## By offense

```{r}
to_plot <- viol_crime_by_off_us |> 
  left_join(offense_pal, by = "crime") |> 
  mutate(
    clearance_rate = clearance_rate * 100,
    crime = fct_reorder(crime, clearance_rate, last),
    tooltip = paste0(
      "<b>", crime, " – ", year, "</b></br>",
      "Clearance rate: ", percent(clearance_rate / 100, 1), "<br>",
      "Offenses reported: ", comma(actual_est, 1), "<br>",
      "Offenses cleared: ", comma(cleared_est, 1)
      )
    ) |> 
  arrange(desc(year), clearance_rate)

to_plot |>   
  hchart(
    "spline",
    hcaes(year, clearance_rate, group = crime),
  ) |> 
  hc_colors(unique(to_plot$color)) |> 
  hc_title(text = paste0("Violent crime clearance rate in ", state_name_text, ", by offense")) |>
  hc_caption(text = "FBI <em>Crime in the U.S.</em>") |> 
  hc_setup() |> 
  hc_yAxis(labels = list(format = "{value:,.0f}%"))
```

:::

```{r, eval=FALSE}
#| column: page-inset
#| out.width: "990px"
#| out.height: "600px"

### Clearance rate of violent crime, 2020

hex <- sf::read_sf("us_states_hexgrid.geojson") |> 
  select(state_abb = iso3166_2)

hex_gj <- hex %>% 
  sf::st_transform(3857) %>%
  geojsonsf::sf_geojson() %>% 
  jsonlite::fromJSON(simplifyVector = FALSE)

highchart_hex_map <- function(x) {
  highchart()  |> 
    hc_add_series_map(
      map = hex_gj,
      df = x,
      joinBy = "state_abb",
      value = "rate",
      dataLabels = list(enabled = TRUE, format = "{point.state_abb}",
                        style = list(fontSize = "12px", fontWeight = "regular",
                                     textOutline = 0)),
      nullColor = "#e8e8e8"
      )  |> 
    hc_colorAxis(minColor = "#f7fdff", maxColor = "#1795BF") |> 
    hc_setup()
}

to_plot <- viol_crime_by_state |> 
  filter(year == 2020) |> 
  mutate(
    clearance_rate = if_else(state_abb == "IL", NA_real_, clearance_rate),
    clearance_rate = clearance_rate * 100,
    rate = clearance_rate,
    tooltip = paste0(
      "<b>", state, " – ", year, "</b><br>",
      "Clearance rate: ", percent(clearance_rate / 100, 1), "<br>",
      "Violent offenses reported: ", comma(actual_est, 1), "<br>",
      "Violent offenses cleared: ", comma(cleared_est, 1)
      ),
    url = paste0("state-viol-crime-", tolower(state_abb), ".html")
    )

highchart_hex_map(to_plot) |> 
  hc_tooltip(enabled = FALSE) |> 
  hc_plotOptions(
    series = list(
      point = list(
        events = list(
          click = JS("function(){location.href = this.options.url}"))
        )
      )
    )
```

## Law enforcement agencies not reporting to NIBRS

Not all law enforcement agencies have begun reporting NIBRS data to the FBI. As of data year 2021, `r comma(yes_nibrs_agencies, 1)` of `r comma(total_agencies, 1)` agencies in `r  state_name_text` reported NIBRS data. Agencies that *did not* report in 2021 are listed in the table below. The population column is the number of residents of the jurisdiction covered by the law enforcement agency.[^zero]

[^zero]: Some agencies, such as university and park police, are considered to have a population of zero by the FBI."


```{r}
ucr_agency_pop |> 
  filter(!is_nibrs) |> 
  mutate(county_name = if_else(county_name == "Not Specified", "", county_name)) |> 
  select(agency_name, state_abbr, county_name, population) |> 
  arrange(desc(population)) |>
  reactable_template(
    sort_col = list("population" = "desc"),
    columns = list(
      agency_name = colDef(name = "Agency", minWidth = 250),
      state_abbr = colDef(name = "State"),
      county_name = colDef(name = "County"),
      population = colDef(name = "Population", style = list(
        fontFamily = "Source Sans Pro",
        fontWeight = 600)))
  )
```
