---
title: "Prep analytic files for violent crime/clearance website updates"
format: html
---

In this file, we will create the following analytic files for the website updates to be rolled out in 1/2025:

* 1. Clearance by type of violent crime (homicide, rape, robbery, agg assault) by LEA


* 2. Resources

  + Budget for LEAs over time – 5-10 years (by state: local and State Police Expenditure)
  
  + Staffing for LEAs over time – 5-10 years
  
  
* 3. Supplemental homicide report data

**Descriptives:**

*	Incidents: clearances, type of clearances, offender and victim counts

*	Offense: location, weapons

*	Victim: relationship with offender, age, gender, race, ethnicity

*	Arrestee: age, gender, race, ethnicity 

*	Highlight characteristics that make cases difficult to solve (e.g., firearms, locations)


**Disparities**

*	In victimization

*	In clearances

*	In case characteristics (e.g., incidents, offense, victim, arrestee)

<br>


```{r }
### call libraries
library(tidyverse)
library(csgjcr)
library(janitor)
library(tidycensus)
library(readxl)
library(sf)

```

```{r }
### set path for clearance rates folder -- where some of the data files are already stored
sp_clearance_rate_data_path <- csg_sp_path("JR_Clearance_Rates/Data")

### set path for main jri data library folder
sp_jr_data_library <- csg_sp_path("JR_DATA_LIBRARY")

### set project path for jr data library for data we want to import
sp_data_analysis_path <- csg_sp_path("JR_DATA_LIBRARY/data/analysis")

### set project path for ad hoc request folder for data we'll import
sp_data_to_plot_path <- csg_sp_path("ad_hoc_requests/state_violent_crime_marshall/data")

### set project path for ad hoc request folder for data we'll export
sp_data_site_updates_path <- csg_sp_path("ad_hoc_requests/state_violent_crime_marshall/data/website_updates")

### set project path for lea crosswalk downloaded from icpsr
sharepoint_jr_clearance_le_agency_crosswalk <- csg_sp_path("JR_Clearance_Rates/Data/Law Enforcement Crosswalk")

### set clean data sharepoint path (for export)
sharepoint_jr_clearance_ucr_clean <- file.path(sp_clearance_rate_data_path,"UCR Data/Clean")

### set path for shr data (through 2022)
sharepoint_jr_clearance_ucr_shr <- file.path(sp_clearance_rate_data_path,"UCR Data/Supplemental Homicide Reports")

### set clean data sharepoint path (for export)
sharepoint_jr_clearance_ucr_employee <- file.path(sp_clearance_rate_data_path,"UCR Data/Employee Data")

```

```{r }
### import nibrs file that matt developed for 50 state snapshots
### file is unique by lea, index crime, and year
### to use for state-specific pages, we need to do the following
# add county (and handle agencies whose jurisdictions include multiple counties? remove? divide by count of counties?)
# create county-level violent index unsolved rate (both 2022 and 2019-2022 change)
df_nibrs_kaplan <- read_rds(file.path(sp_data_analysis_path,"fbi","kaplan","fbi_kaplan_okca_agency.rds")) %>% 
  janitor::clean_names() |> 
  ### filter out agencies without all 12 months of data
  filter(number_of_months_reported==12)


### pull in list of states and DC to filter out territories in other files
states_list <- csgjcr::state_fips_lookup |> 
  filter(state_fips<=56) 

```

```{r }
### next we will pull in a law enforcement agency lookup table in order to add a few more columns
### namely, we want county and cleaned up agency name 


### import fbi law enforcement agency crosswalk from icpsr download
### downloaded from: https://www.icpsr.umich.edu/web/NACJD/studies/35158#
### we'll use this crosswalk to join county to each agency

### import 2012 raw le_agency_crosswalk.rda from icpsr download
load(file=file.path(sharepoint_jr_clearance_le_agency_crosswalk, "le_agency_crosswalk.rda"))

# in 2022 ct change the names and boundaries of their counties
# this will upate the existing crosswalk with a new fips for each ori
# going to upate all the ct agencies with these new fips
ct_county_update <- read_csv(file.path(sp_data_to_plot_path,"ct-county-update-ori.csv")) |> 
  dplyr::select(ori, county_fips = new_county_fips)

### rename and clean dataframe
### note: there are ~1,000 agencies who are missing ORI IDs (ori == -1)
### these largely appear to be small and unconvential PDs (like airport or state park departments)
### we don't explicitly drop them here, but they'll be dropped when we left join this file to our analytic file
### since we join on ORI ID
df_le_agency_crosswalk <- da35158.0001 |> 
  clean_names() |> 
  ### concatenate fips_state_code and fips_county_code 
  ### clean names
  mutate(county_fips = paste0(fips_st,
                              fips_county),
         county = trimws(str_to_title(countyname),
                         which = "both"),
         state_name = trimws(str_to_title(as.character(statename)),
                             which = "both"),
         name = str_to_title(name)) |> 
  dplyr::select(agency_name_lookup = name,
                ori = ori9,
                county,
                county_fips) |> 
  ### NOTE: SOME AGENCIES APPEAR TO HAVE WRONG ORI ID
  ### BIGGEST EXAMPLE IS PHILADELPHIA PD
  ### BASED ON ORI IN THIS CROSSWALK
  ### WE JOIN TO INDIANA COUNTY INSTEAD
  ### OF PHILADELPHIA COUNTY
  ### hard code fix for philly
  ### is this just due to the 2012 file being outdated?
  mutate(county = if_else(ori=="PAPEP0000",
                      "Philadelphia",county),
         county_fips = if_else(ori=="PAPEP0000",
                      "42101",county_fips)) |>
  rows_update(ct_county_update, by = "ori")


### remove original file
rm(da35158.0001)

```

```{r }
################################################################
### pull in county-level population estimates
### we end up just using this file for cleaned up county names
### they appear to be better than the names from UCR/NIBRS
################################################################

### county-level file
### pulling in population data in case we want to use it as tooltip
us_county_population <- get_acs(
  geography = "county",
  variables = "S0101_C01_001",
  year = 2022,
  geometry = TRUE,
  keep_geo_vars = TRUE,
  cb = TRUE,
  resolution = "20m") |>
  clean_names() |>
  mutate(state_geoid = substr(geoid, 1, 2)) |>
  ### now drop geometry so that we can use as regular df (and not a spatial object)
  ### we initially needed geometry in order to extract county name without label (e.g., county, parish, area, etc.)
  st_drop_geometry() |> 
  ### finally, keep and rename columns we'll use later
  dplyr::select(county_fips = geoid,
                acs_county_name = name_x,
                acs_county_pop = estimate,
                acs_county_state_name = state_name) |>
  distinct() 

```

## 1a. Clearance by type of violent crime (homicide, rape, robbery, agg assault) by LEA

```{r }
###################################################
### this chunk produces an analytic file
### for the updated section of the site
### with clearance rates by violent crime type
### and % change over last 5 years (2018-2023)
### this syntax is similar to above except that 
### we are additionally grouping by "group" column
###################################################


###################################################
### note: 'group' = type of offense
### create agency- and group-level file for table
### two measures: 2023 unsolved rate 
### and 2018-2023 percent change
#################################################
df_nibrs_violent_unsolved_by_type_year_agency <- df_nibrs_kaplan |> 
  ### keep offense types we want in final analysis
  ### and only keep agencies that have reported data for all 12 months
  filter(group %in% c("Homicide", "Rape", "Robbery", "Aggravated assault"),
         number_of_months_reported==12) |> 
  ### add county from lookup
  left_join(df_le_agency_crosswalk,
            by = "ori") |> 
  ### calculate rates
  mutate(unsolved_rate = n_unsolved/n_reported,
         n_reported_rate = n_reported/pop_jurisdiction*100000) |> 
  ### recode any unsolved counts or rates less than 0 to 0
  mutate(unsolved_rate = ifelse(unsolved_rate<0 & !is.na(unsolved_rate),
                                     0,unsolved_rate),
         n_unsolved = ifelse(n_unsolved<0 & !is.na(n_unsolved),
                                     0,n_unsolved)) 


### prep df for agency-level table by type of offense
df_nibrs_violent_unsolved_by_agency_type_to_table <- df_nibrs_violent_unsolved_by_type_year_agency |> 
  ### now calculate solve rate instead of unsolved rate
  ### by backing into numerator and denominator
  mutate(n_unsolved = unsolved_rate*n_reported,
         n_solved = n_reported - n_unsolved,
         solved_rate = n_solved / n_reported)  |>
  dplyr::select(-n_unsolved) |> 
  dplyr::select(year,
                  agency_name,
                  agency_type,
                  ori,
                  county,
                  county_fips,
                  state_name,
                  state_abbr,
                  population = pop_jurisdiction,
                  group,
                  n_reported:solved_rate, 
                  number_of_months_reported)

### first df with everything we need except for calculation of 2019 to 2023 change
df_nibrs_violent_unsolved_by_agency_type_to_table_2023_1 <- df_nibrs_violent_unsolved_by_agency_type_to_table |> 
  filter(year==2023)

### now calculate 2019-2023 change where possible and join to df_nibrs_violent_unsolved_by_agency_to_table_2023_1
### agency must have both years of data for this calculation
df_nibrs_violent_unsolved_by_agency_type_to_table_2023_to_join <- df_nibrs_violent_unsolved_by_agency_type_to_table |> 
  filter(year%in%c(
    2018,
    2023)) |>
  group_by(ori,
           group) |>
  mutate(years_of_data = n_distinct(year)) |>
  filter(years_of_data==2) |>
  ### currently multiplying rate by 100 instead of using scales::percent
  ### may be better way to format within hc
  mutate(solved_rate_2018 = solved_rate[year==2018],
         solved_rate_2023 = solved_rate[year==2023],
         # solved_rate_percent_change_2019_2023 = ((solved_rate_2023 - solved_rate_2019)/solved_rate_2019),
         solved_rate_percent_point_change_2018_2023 = solved_rate_2023 - solved_rate_2018,
         ### convert to percent point (from decimal)
         solved_rate_percent_point_change_2018_2023 = solved_rate_percent_point_change_2018_2023*100) |>
  ungroup() |>
  ### now drop 2018
  ### and keep unsolved rate from latest year
  filter(year==2023) |> 
    dplyr::select(ori,
                  group,
                  solved_rate_percent_point_change_2018_2023) |> 
  mutate(solved_rate_percent_point_change_2018_2023 = ifelse(is.nan(solved_rate_percent_point_change_2018_2023),
                                                       NA,solved_rate_percent_point_change_2018_2023))


### now join df_nibrs_violent_unsolved_by_agency_to_table_2023 and df_nibrs_violent_unsolved_by_agency_to_table_2023_to_join
df_nibrs_violent_unsolved_by_agency_and_type_to_table_2023 <- left_join(df_nibrs_violent_unsolved_by_agency_type_to_table_2023_1,
                                                               df_nibrs_violent_unsolved_by_agency_type_to_table_2023_to_join,
                                                               by = c("ori",
                                                                      "group")) |> 
  ### pivot long to wide
  ### one row per agency
  pivot_wider(id_cols = c(year,
                  agency_name,
                  agency_type,
                  ori,
                  county,
                  county_fips,
                  state_name,
                  state_abbr,
                  population,
                  number_of_months_reported),
              names_from = group, 
              values_from = c(n_reported,n_reported_rate,n_solved,solved_rate,solved_rate_percent_point_change_2018_2023)) |> 
  ### clean up column names from pivot
  clean_names() |> 
  ### join acs county level file to clean up county name
  ### similar to mapping df used for site
  left_join(us_county_population,
            by = "county_fips") |>
  ### county name from ACS appears cleaner than from lea lookup table
  ### so we'll clean up `county` -- using the acs version when not missing from lookup 
  ### using coalesce: https://dplyr.tidyverse.org/reference/coalesce.html
  mutate(county_clean = coalesce(acs_county_name,county)) |> 
  dplyr::select(agency_name,
                  agency_type,
                  ori,
                  county=county_clean,
                  state_name,
                  state_abbr,
                  jurisdiction_population = population,
                  n_reported_aggravated_assault:solved_rate_percent_point_change_2018_2023_robbery, 
                  number_of_months_reported) 



### write out
write_rds(df_nibrs_violent_unsolved_by_agency_and_type_to_table_2023,
          file.path(sp_data_site_updates_path,"website_update_viol_crime_unsolved_by_agency_and_type.rds"))

```

<br>

## 1b. Clearance by type of violent crime (homicide, rape, robbery, agg assault) by state

Note: This is for the national page.

```{r }
###################################################
### this chunk produces an analytic file
### for the updated national page of the site
### with clearance rates by violent crime type
### for latest year available
###################################################

### read in crime data
### using srs_state to align numbers/methods with what is already on state pages
### for incident rates and solve rates
srs_state <- read_rds(file.path(sp_data_to_plot_path, "srs_state.rds"))

srs_state_by_offense <- srs_state |> 
  filter(crime_cat == "Violent", 
         year == 2023) |>
  mutate(
    n_solved = incidents_reported - incidents_unsolved,
    solved_rate = n_solved / incidents_reported,
    n_reported_rate = incidents_reported/pop_total*100000
  ) |> 
  ### recode any unsolved counts or rates greater than 1 to 1
  mutate(solved_rate = ifelse(solved_rate>1 & !is.na(solved_rate),
                                     1,solved_rate)) |> 
  ### pivot long to wide
  ### one row per agency
  pivot_wider(id_cols = c(state_name),
              names_from = group,
              values_from = c(n_reported_rate,
                              solved_rate)) |>
  ### clean up column names from pivot
  clean_names() |>
  dplyr::select(state_name,
                n_reported_rate_aggravated_assault:solved_rate_robbery)


### write out
write_rds(srs_state_by_offense,
          file.path(sp_data_site_updates_path,"website_update_viol_crime_unsolved_by_state_and_type.rds"))


```

<br><br>

## 2. Resources

### Budget for LEAs over time – 5-10 years (by state: local and State Police Expenditure)

**Note:** All expenditure data has been adjusted to account for income inflation over time using the following formula (e.g. to adjust 1995 data to 2022 numbers):

* 1995 Inflation-adjusted estimate = 1995 estimate * (2022 index value / 1995 index value)

See: https://www.census.gov/topics/income-poverty/income/guidance/current-vs-constant-dollars.html


**Source:** https://www.census.gov/programs-surveys/gov-finances/data/datasets.All.List_1883146942.html#list-tab-List_1883146942

```{r include=FALSE,message=FALSE,warning=FALSE}

### assign year to earliest year to keep value for filter in code below
earliest_year_to_keep <- 2012

### import census survey of local and state gov expenditure
asslgf <- read_rds(file.path(sp_jr_data_library,"data","analysis","census", "asslgf", "census_asslgf_state.rds")) 

### import inflation index info for calculation of inflation-adjusted expenses
#### from here: https://www.census.gov/topics/income-poverty/income/guidance/current-vs-constant-dollars.html
inflation_index_years <- read_xls(file.path(sp_clearance_rate_data_path,"annual-index-value_annual-percent-change.xls")) |> 
  clean_names() |> 
  dplyr::select(year=income_year,
                inflation_index = index)

### extract inflation index for most recent year
inflation_index_latest <- inflation_index_years$inflation_index[inflation_index_years$year==max(inflation_index_years$year)]

### adjust expenditure data to be inflation adjusted
### state-level
asslgf_clean <- asslgf |> 
  filter(indicator=="Expenditures",
         group=="Police Protection") |> 
  left_join(inflation_index_years,
            by = c("year")) |> 
  ### resource for adjusting for inflation: https://www.census.gov/topics/income-poverty/income/guidance/current-vs-constant-dollars.html
  mutate(n_inflation_adjusted_latest_year_dollars = n * (inflation_index_latest / inflation_index),
         inflation_adjusted_rate_total = n_inflation_adjusted_latest_year_dollars/pop_total) |> 
  ### keep last 10 years of data
  filter(year>=earliest_year_to_keep)


### adjust expenditure data to be inflation adjusted
### national-level
asslgf_clean_national <- asslgf_clean |> 
  ### now sum dollars across state by year for national page viz df
  group_by(year) |> 
  summarize(n_inflation_adjusted_latest_year_dollars = sum(n_inflation_adjusted_latest_year_dollars,
                                                    na.rm=TRUE)) |> 
  ungroup()

### write out

## state df
write_rds(asslgf_clean,
          file.path(sp_data_site_updates_path,"website_update_state_level_police_expenditure_inflation_adjusted.rds"))

## national df
write_rds(asslgf_clean_national,
          file.path(sp_data_site_updates_path,"website_update_national_level_police_expenditure_inflation_adjusted.rds"))

```

<br>

#### Local and State Police Expenditure as % of Total Budget

```{r include=FALSE,message=FALSE,warning=FALSE}

### now look at police spending as % of total budget
asslgf_percent_clean <- asslgf |> 
  filter(indicator=="Percent of total state expenditure",
         group=="Police Protection") |> 
  ### keep last 10 years of data
  filter(year>=earliest_year_to_keep)

### write out
write_rds(asslgf_percent_clean,
          file.path(sp_data_site_updates_path,"website_update_state_level_police_expenditure_percent_of_budget.rds"))

```

### Staffing for LEAs over time – 5-10 years


#### UCR: law enforcement employee data 

**Note:** In the following analysis, I'm using the count of sworn officers only, not civilians. Also, Nebraska is removed because they don't appear to have any reported lea employee data in this sample window

**Source:**  https://cde.ucr.cjis.gov/LATEST/webapp/#/pages/downloads

```{r }
### import and clean file
### calculate officers per 10k residents in agency jurisdiction
ucr_employee_data_raw <- readr::read_csv(file.path(sharepoint_jr_clearance_ucr_employee,
                                         "lee_1960_2023.csv")) |> 
  dplyr::select(ori,
                year=data_year,
                state_abbr,
                agency_type_name,
                population,
                officer_ct, 
                civilian_ct) |> 
  filter(year>=earliest_year_to_keep) |> 
  mutate(officer_pop_rate_per_10k = (officer_ct/population)*10000,
         state_name = csgjcr::csg_state_convert(state_abbr,
                                                from = "abbr",
                                                to = "name")) 

### calculate officers per 10k residents nationally
ucr_employee_data_raw_national <- ucr_employee_data_raw |>
  group_by(year) |> 
  summarize(officer_ct = sum(officer_ct,
                             na.rm=TRUE),
            population = sum(population,
                             na.rm=TRUE),
            officer_pop_rate_per_10k = (officer_ct/population)*10000) |> 
  ungroup()

```

<br>


### Workload over time – 5-10 years

*	Officers per capita

*	Officers per violent crime

```{r }
########################################
### create agency-level file for join
### to employee data
########################################

## agency-level
df_nibrs_violent_unsolved_by_year_agency_offense_type <- df_nibrs_kaplan |> 
  filter(group %in% c("Homicide", "Rape", "Robbery", "Aggravated assault"),
         number_of_months_reported==12,
         year >= earliest_year_to_keep) |> 
  dplyr::group_by(year,
                  ori) |> 
  summarize(n_all_violent_reported = sum(n_reported,
                             na.rm = TRUE)) |> 
  ungroup() 

### now calculate officers per violent crime and homicide over time
ucr_employee_data_clean <- left_join(ucr_employee_data_raw,
                                     df_nibrs_violent_unsolved_by_year_agency_offense_type,
                                     by = c("ori",
                                            "year")) |> 
  mutate(officer_per_violent_offense_1k = (officer_ct/n_all_violent_reported)*1000)


## national-level
## use national violent crime file (srs_state)
srs_by_cat_us <- srs_state |>
  filter(crime_cat == "Violent", 
         year >= earliest_year_to_keep) |>
  group_by(year) |>
  summarize(
    across(c(pop_total, incidents_reported, incidents_unsolved), \(x) sum(x, na.rm = TRUE))
  ) |>
  dplyr::select(year,
                n_all_violent_reported = incidents_reported)


### now calculate officers per violent crime and homicide over time
ucr_employee_data_clean_national <- left_join(ucr_employee_data_raw_national,
                                     srs_by_cat_us,
                                     by = c("year")) |> 
  mutate(officer_per_violent_offense_1k = (officer_ct/n_all_violent_reported)*1000)


########################################
### now create state-level file for join
### to employee data
########################################

### import and clean file at state-level
### calculate officers per 10k residents in agency jurisdiction
ucr_employee_data_raw_state <- ucr_employee_data_raw |> 
  group_by(state_name,
           state_abbr,
           year) |> 
  summarise(officer_ct = sum(officer_ct,na.rm=TRUE),
            population = sum(population,na.rm=TRUE),
            officer_pop_rate_per_10k = (officer_ct/population)*10000) |> 
  ungroup()

### prep offense data at state level for join
df_nibrs_violent_unsolved_by_year_agency_offense_type_state <- df_nibrs_kaplan |> 
  filter(group %in% c("Homicide", "Rape", "Robbery", "Aggravated assault"),
         number_of_months_reported==12,
         year >= 2012) |> 
  dplyr::group_by(state_abbr,
                  year) |> 
  summarize(n_all_violent_reported = sum(n_reported,
                             na.rm = TRUE),
            n_cleared = sum(n_cleared,
                             na.rm = TRUE),
            n_homicide_reported = sum(n_reported[group=="Homicide"])) |> 
  ungroup() |> 
  mutate(solve_rate = n_cleared/n_all_violent_reported)

### now calculate officers per violent crime and homicide over time
### by state
ucr_employee_data_by_state_clean <- left_join(ucr_employee_data_raw_state,
                                     df_nibrs_violent_unsolved_by_year_agency_offense_type_state,
                                     by = c("state_abbr",
                                            "year")) |> 
  mutate(officer_per_violent_offense_1k = (officer_ct/n_all_violent_reported)*1000,
         officer_per_homicide_100 = (officer_ct/n_homicide_reported)*100)



### write out

## national
write_rds(ucr_employee_data_clean_national,
          file.path(sp_data_site_updates_path,"website_update_lea_staffing_workload_national.rds"))

## agency level
write_rds(ucr_employee_data_clean,
          file.path(sp_data_site_updates_path,"website_update_lea_staffing_workload.rds"))

### state level
write_rds(ucr_employee_data_by_state_clean,
          file.path(sp_data_site_updates_path,"website_update_lea_staffing_workload_by_state.rds"))

```


<br><br>

## 3. Supplemental homicide report data 

### Descriptives

*	Incidents: clearances, type of clearances, offender and victim counts

*	Offense: location, weapons

*	Victim: relationship with offender, age, gender, race, ethnicity

*	Arrestee: age, gender, race, ethnicity 

*	Highlight characteristics that make cases difficult to solve (e.g., firearms, locations)

<br>

### Disparities (also supplemental homicide report data)

*	In victimization

*	In clearances

*	In case characteristics (e.g., incidents, offense, victim, arrestee)

**Methodological notes:** include the % of homicides represented in the data.  Should this be giant table by LEA or aggregated to county or state levels?

**NOTE:** Any disaggregation by a variable that is correlated with a clearance (e.g., arrestee demographics, weapon used, victim/arresteee relationship, etc.) should be interpreted with caution and included with a disclaimer. 


```{r include=FALSE,message=FALSE,warning=FALSE}
### read in crime data
ucr_srs_state <- read_rds(file.path(sp_data_to_plot_path, "srs_state.rds")) |> 
  filter(group == "Homicide", 
         year == 2022) |> 
  dplyr::select(state_name,
                state_abb = state_abbr,
                incidents_reported)
  
```

```{r include=FALSE,message=FALSE,warning=FALSE}
### read in shr_1976_2022
### downloaded from icprsr: https://www.openicpsr.org/openicpsr/project/100699/version/V14/view?path=/openicpsr/100699/fcr:versions/V14/shr_1976_2022.rds&type=file
shr_1976_2022 <- read_rds(file.path(sharepoint_jr_clearance_ucr_shr, "shr_1976_2022.rds")) |> 
  ### clean up grouping vars we'll use in the next chunk
  mutate(victim_1_relation_to_offender_1 = str_to_title(victim_1_relation_to_offender_1),
         offender_1_circumstance = str_to_title(offender_1_circumstance),
         offender_1_weapon = str_to_title(offender_1_weapon),
         ### clean up weapon type
         offender_1_weapon_clean = ifelse(offender_1_weapon%in%c("Handgun","Firearm, Type Not Stated", "Rifle", "Other Gun", "Shotgun"), 
                                          "Firearm",offender_1_weapon),
         offender_1_weapon_clean = ifelse(offender_1_weapon=="Narcotics/Drugs - Includes Sleeping Pills", 
                                          "Narcotics/Drugs",offender_1_weapon_clean),
  ### clean race/ethnicity and gender
  ### use factor function to assign labels to values
  victim_1_race_clean = case_when(
    (victim_1_race=="american indian or alaskan native") & victim_1_ethnic_origin!="hispanic" ~ 1,
    (victim_1_race%in%c("asian","native hawaiian or other pacific islander")) & victim_1_ethnic_origin!="hispanic" ~ 2,
    (victim_1_race=="black") & victim_1_ethnic_origin!="hispanic" ~ 3,
    victim_1_ethnic_origin=="hispanic" ~ 4,
    (victim_1_race=="white") & victim_1_ethnic_origin!="hispanic" ~ 5,
    victim_1_race=="unknown" & victim_1_ethnic_origin!="hispanic" ~ 6,
    TRUE ~ 7),
    victim_1_race_clean = factor(victim_1_race_clean,
         levels = c(1,2,3,4,5,6,7),
         labels = c("American Indian/Alaskan Native",
                    "Asian/Pacific Islander",
                    "Black",
                    "Hispanic",
                    "White",
                    "Unknown",
                    "Missing")),
  victim_1_sex_clean = case_when(
    victim_1_sex=="female" ~ 1,
    victim_1_sex=="male" ~ 2,
    victim_1_sex=="unknown" ~ 3,
    TRUE ~ as.numeric(NA)),
    victim_1_sex_clean = factor(victim_1_sex_clean,
         levels = c(1,2,3),
         labels = c("Female", "Male", "Unknown")),
  ### create age bins
  ### there appear to be juveniles in data so naming lowest age category as under 24
  ### exclude ages over 105 -- these are likely data entry errors and should be coded as N/A
  ### for this cleaned up variable (they'll be excluded in viz)
  victim_1_age = as.numeric(victim_1_age),
  victim_1_age_bin = case_when(
    victim_1_age<25 ~ 1,
    victim_1_age>=25 & victim_1_age<=34 ~ 2,
    victim_1_age>34 & victim_1_age<=45 ~ 3,
    victim_1_age>45 & victim_1_age<=105 ~ 4,
    TRUE ~ as.numeric(NA)),
    victim_1_age_bin = factor(victim_1_age_bin,
         levels = c(1,2,3,4),
         labels = c("Under 25", "25 to 34", "35 to 45", "46+")),
  ### defining multiple victims as instances where victim_2 has information populated for any demographics
  victim_multiple_flag = ifelse((!is.na(victim_2_age)|!is.na(victim_2_ethnic_origin)|!is.na(victim_2_race)|!is.na(victim_2_sex)),
                                       "Multiple Victims","Single Victim")) |> 
  dplyr::rename(circumstance_of_homicide = offender_1_circumstance) 


### clean up shr
### overall
shr_2022_state <- shr_1976_2022 |> 
  filter(year==2022) |> 
  ### calculate clearance rate
  ### code any incident where offender information is populated (e.g. age, sex, or race/ethnicity) as cleared
  mutate(cleared_flag = ifelse(offender_1_age!="unknown" & !is.na(offender_1_age) & 
                                 offender_1_race!="unknown" & !is.na(offender_1_race) &
                                 offender_1_sex!="unknown" & !is.na(offender_1_sex),
                               1,0),
         cleared_flag = ifelse(is.na(cleared_flag),
                               0,cleared_flag)) |> 
  group_by(state_abb,
           year) |> 
  summarize(n_total_incidents = n(),
            n_total_cleared = sum(cleared_flag,na.rm = TRUE),
            clearance_rate = n_total_cleared/n_total_incidents) |> 
  ungroup() 

### join shr and ucr data to explore the percent of all ucr reported homicides that appear in SHR
shr_2022_state_ucr <- shr_2022_state |> 
  left_join(ucr_srs_state,
            by = "state_abb") |> 
  mutate(`Percent Coverage (SHR/UCR SRS)` = scales::percent(n_total_incidents/incidents_reported,
                                                            accuracy = 1)) |> 
  dplyr::select(`State` = state_name,
                `Percent Coverage (SHR/UCR SRS)`,
                `SHR Homicides (N)` = n_total_incidents,
                `UCR SRS Homicides (N)` = incidents_reported) |> 
  filter(!is.na(`State`))


### write out
write_rds(shr_2022_state_ucr,
          file.path(sp_data_site_updates_path,
                    "shr_reported_homicides_ucr_srs_coverage_2022.xlsx"))

```


```{r }

### overall
shr_1976_2022_national_overall <- shr_1976_2022 |> 
  ### calculate clearance rate
  ### code any incident where offender information is populated (e.g. age, sex, or race/ethnicity) as cleared
  mutate(cleared_flag = ifelse(offender_1_age!="unknown" & !is.na(offender_1_age) & 
                                 offender_1_race!="unknown" & !is.na(offender_1_race) &
                                 offender_1_sex!="unknown" & !is.na(offender_1_sex),
                               1,0),
         cleared_flag = ifelse(is.na(cleared_flag),
                               0,cleared_flag)) |> 
  group_by(year) |> 
  summarize(n_total_incidents = n(),
            n_total_cleared = sum(cleared_flag,na.rm = TRUE),
            clearance_rate = n_total_cleared/n_total_incidents) |> 
  ungroup()

# ### double check coding for age
# table(shr_1976_2022_by_agency_over_time$offender_1_age[shr_1976_2022_by_agency_over_time$cleared_flag==0], useNA = 'always')
# 
# ### double check coding for race
# table(shr_1976_2022_by_agency_over_time$offender_1_race[shr_1976_2022_by_agency_over_time$cleared_flag==0])
# 
# ### double check coding for sex
# table(shr_1976_2022_by_agency_over_time$offender_1_sex[shr_1976_2022_by_agency_over_time$cleared_flag==0])

```


```{r }

### create function to clean up shr for national page viz
function_clean_shr_1976_2022_national_by_group <- function(df, grouping_var){
  
  df |> 
  ### calculate clearance rate
  ### code any incident where offender information is populated (e.g. age, sex, or race/ethnicity) as cleared
  mutate(cleared_flag = ifelse(offender_1_age!="unknown" & !is.na(offender_1_age) & 
                                 offender_1_race!="unknown" & !is.na(offender_1_race) &
                                 offender_1_sex!="unknown" & !is.na(offender_1_sex),
                               1,0),
         cleared_flag = ifelse(is.na(cleared_flag),
                               0,cleared_flag)) |> 
  group_by(year,
           !!sym(grouping_var)) |> 
  summarize(n_total_incidents = n(),
            n_total_cleared = sum(cleared_flag,na.rm = TRUE),
            clearance_rate = n_total_cleared/n_total_incidents) |> 
  ungroup() 
  
}


### now create files ready for viz with function

### relationship
shr_1976_2022_national_relationship <- function_clean_shr_1976_2022_national_by_group(df = shr_1976_2022,
                                                                                      grouping_var = "victim_1_relation_to_offender_1")

### weapon used
shr_1976_2022_national_weapon <- function_clean_shr_1976_2022_national_by_group(df = shr_1976_2022,
                                                                                      grouping_var = "offender_1_weapon_clean")
  
### circumstance of homicide
shr_1976_2022_national_circumstance <- function_clean_shr_1976_2022_national_by_group(df = shr_1976_2022,
                                                                                      grouping_var = "circumstance_of_homicide")

### race
shr_1976_2022_national_race <- function_clean_shr_1976_2022_national_by_group(df = shr_1976_2022,
                                                                                      grouping_var = "victim_1_race_clean")

### gender
shr_1976_2022_national_gender <- function_clean_shr_1976_2022_national_by_group(df = shr_1976_2022,
                                                                                      grouping_var = "victim_1_sex_clean")

### age
shr_1976_2022_national_age <- function_clean_shr_1976_2022_national_by_group(df = shr_1976_2022,
                                                                                      grouping_var = "victim_1_age_bin") |> 
  ### remove rows where age is NA
  filter(!is.na(victim_1_age_bin))

```

```{r }
### now write all dfs out for visualization

### write out
write_rds(shr_1976_2022_national_overall,
          file.path(sp_data_site_updates_path,"shr_1976_2022_national_overall_clear.rds"))

### write out
write_rds(shr_1976_2022_national_relationship,
          file.path(sp_data_site_updates_path,"shr_1976_2022_national_victim_offender_relationship.rds"))

### write out
write_rds(shr_1976_2022_national_weapon,
          file.path(sp_data_site_updates_path,"shr_1976_2022_national_weapon.rds"))

### write out
write_rds(shr_1976_2022_national_circumstance,
          file.path(sp_data_site_updates_path,"shr_1976_2022_national_incident_circumstance.rds"))

### write out
write_rds(shr_1976_2022_national_race,
          file.path(sp_data_site_updates_path,"shr_1976_2022_national_victim_race_eth.rds"))

### write out
write_rds(shr_1976_2022_national_gender,
          file.path(sp_data_site_updates_path,"shr_1976_2022_national_victim_gender.rds"))

### write out
write_rds(shr_1976_2022_national_age,
          file.path(sp_data_site_updates_path,"shr_1976_2022_national_victim_age.rds"))

```

<br>

```{r }
########################
### now do the same for 
### state-level dfs
########################

### overall
shr_1976_2022_state_overall <- shr_1976_2022 |> 
  dplyr::rename(state_abbr=state_abb) |> 
  ### add state name for eventual viz, as well
  mutate(state_name = csgjcr::csg_state_convert(state_abbr,
                                                from = "abbr",
                                                to = "name")) |> 
  ### calculate clearance rate
  ### code any incident where offender information is populated (e.g. age, sex, or race/ethnicity) as cleared
  mutate(cleared_flag = ifelse(offender_1_age!="unknown" & !is.na(offender_1_age) & 
                                 offender_1_race!="unknown" & !is.na(offender_1_race) &
                                 offender_1_sex!="unknown" & !is.na(offender_1_sex),
                               1,0),
         cleared_flag = ifelse(is.na(cleared_flag),
                               0,cleared_flag)) |> 
  group_by(state_name,
           state_abbr,
           year) |> 
  summarize(n_total_incidents = n(),
            n_total_cleared = sum(cleared_flag,na.rm = TRUE),
            clearance_rate = n_total_cleared/n_total_incidents) |> 
  ungroup() 

```



```{r }

### create function to clean up shr for state page viz
function_clean_shr_1976_2022_state_by_group <- function(df, grouping_var){
  
  df |> 
  dplyr::rename(state_abbr=state_abb) |> 
  ### add state name for eventual viz, as well
  mutate(state_name = csgjcr::csg_state_convert(state_abbr,
                                                from = "abbr",
                                                to = "name")) |> 
  ### calculate clearance rate
  ### code any incident where offender information is populated (e.g. age, sex, or race/ethnicity) as cleared
  mutate(cleared_flag = ifelse(offender_1_age!="unknown" & !is.na(offender_1_age) & 
                                 offender_1_race!="unknown" & !is.na(offender_1_race) &
                                 offender_1_sex!="unknown" & !is.na(offender_1_sex),
                               1,0),
         cleared_flag = ifelse(is.na(cleared_flag),
                               0,cleared_flag)) |> 
  group_by(state_name,
           state_abbr,
           year,
           !!sym(grouping_var)) |> 
  summarize(n_total_incidents = n(),
            n_total_cleared = sum(cleared_flag,na.rm = TRUE),
            clearance_rate = n_total_cleared/n_total_incidents) |> 
  ungroup() 
  
}


### now create files ready for viz with function

### relationship
shr_1976_2022_state_relationship <- function_clean_shr_1976_2022_state_by_group(df = shr_1976_2022,
                                                                                      grouping_var = "victim_1_relation_to_offender_1")

### weapon used
shr_1976_2022_state_weapon <- function_clean_shr_1976_2022_state_by_group(df = shr_1976_2022,
                                                                                      grouping_var = "offender_1_weapon_clean")
  
### circumstance of homicide
shr_1976_2022_state_circumstance <- function_clean_shr_1976_2022_state_by_group(df = shr_1976_2022,
                                                                                      grouping_var = "circumstance_of_homicide")

### single victim vs multiple victims
shr_1976_2022_state_single_multiple_victims <- function_clean_shr_1976_2022_state_by_group(df = shr_1976_2022,
                                                                                      grouping_var = "victim_multiple_flag")

### race
shr_1976_2022_state_race <- function_clean_shr_1976_2022_state_by_group(df = shr_1976_2022,
                                                                                      grouping_var = "victim_1_race_clean")

### gender
shr_1976_2022_state_gender <- function_clean_shr_1976_2022_state_by_group(df = shr_1976_2022,
                                                                                      grouping_var = "victim_1_sex_clean")

### age
shr_1976_2022_state_age <- function_clean_shr_1976_2022_state_by_group(df = shr_1976_2022,
                                                                                      grouping_var = "victim_1_age_bin") |> 
  ### remove rows where age is NA
  filter(!is.na(victim_1_age_bin))

```


```{r }
### now write all dfs out for visualization

### write out
write_rds(shr_1976_2022_state_overall,
          file.path(sp_data_site_updates_path,"shr_1976_2022_state_overall_clear.rds"))

### write out
write_rds(shr_1976_2022_state_relationship,
          file.path(sp_data_site_updates_path,"shr_1976_2022_state_victim_offender_relationship.rds"))

### write out
write_rds(shr_1976_2022_state_single_multiple_victims,
          file.path(sp_data_site_updates_path,"shr_1976_2022_state_single_multiple_victims.rds"))

### write out
write_rds(shr_1976_2022_state_weapon,
          file.path(sp_data_site_updates_path,"shr_1976_2022_state_weapon.rds"))

### write out
write_rds(shr_1976_2022_state_circumstance,
          file.path(sp_data_site_updates_path,"shr_1976_2022_state_incident_circumstance.rds"))

### write out
write_rds(shr_1976_2022_state_race,
          file.path(sp_data_site_updates_path,"shr_1976_2022_state_victim_race_eth.rds"))

### write out
write_rds(shr_1976_2022_state_gender,
          file.path(sp_data_site_updates_path,"shr_1976_2022_state_victim_gender.rds"))

### write out
write_rds(shr_1976_2022_state_age,
          file.path(sp_data_site_updates_path,"shr_1976_2022_state_victim_age.rds"))

```


<br>

