---
title: "Prep analytic files for violent crime/clearance website updates"
format: html
---

In this file, we will create the following analytic files for the website updates to be rolled out in 1/2025:

* 1. Clearance by type of violent crime (homicide, rape, robbery, agg assault) by LEA


* 2. Resources

  + Budget for LEAs over time – 5-10 years (by state: local and State Police Expenditure)
  
  + Staffing for LEAs over time – 5-10 years
  
  
* 3. Supplemental homicide report data

**Descriptives:**

*	Incidents: clearances, type of clearances, offender and victim counts

*	Offense: location, weapons

*	Victim: relationship with offender, age, gender, race, ethnicity

*	Arrestee: age, gender, race, ethnicity 

*	Highlight characteristics that make cases difficult to solve (e.g., firearms, locations)


**Disparities**

*	In victimization

*	In clearances

*	In case characteristics (e.g., incidents, offense, victim, arrestee)

<br>


```{r }
### call libraries
library(tidyverse)
library(csgjcr)
library(janitor)
library(tidycensus)
library(readxl)
```

```{r }
### set path for use in combination with additional subfolder names
sp_clearance_rate_data_path <- csg_get_project_path("clearance_rates")

### set path for use in combination with additional subfolder names
sp_jr_data_library <- csg_get_project_path("jr_data_library")

### set project path for jr data library for data we want to import
sp_data_analysis_path <- csg_sp_path("JR_DATA_LIBRARY/data/analysis")

### set project path for ad hoc request folder for data we'll import
sp_data_to_plot_path <- csg_sp_path("ad_hoc_requests/state_violent_crime_marshall/data")

### set project path for ad hoc request folder for data we'll export
sp_data_site_updates_path <- csg_sp_path("ad_hoc_requests/state_violent_crime_marshall/data/website_updates")

### set project path for lea crosswalk downloaded from icpsr
sharepoint_jr_clearance_le_agency_crosswalk <- csg_sp_path("JR_Clearance_Rates/Data/Law Enforcement Crosswalk")

```

```{r }
### import nibrs file that matt developed for 50 state snapshots
### file is unique by lea, index crime, and year
### to use for state-specific pages, we need to do the following
# add county (and handle agencies whose jurisdictions include multiple counties? remove? divide by count of counties?)
# create county-level violent index unsolved rate (both 2022 and 2019-2022 change)
df_nibrs_kaplan <- read_rds(file.path(sp_data_analysis_path,"fbi","kaplan","fbi_kaplan_okca_agency.rds")) %>% 
  janitor::clean_names() |> 
  ### filter out agencies without all 12 months of data
  filter(number_of_months_reported==12)


### pull in list of states and DC to filter out territories in other files
states_list <- csgjcr::state_fips_lookup |> 
  filter(state_fips<=56) 

```

```{r }
### next we will pull in a law enforcement agency lookup table in order to add a few more columns
### namely, we want county and cleaned up agency name 


### import fbi law enforcement agency crosswalk from icpsr download
### downloaded from: https://www.icpsr.umich.edu/web/NACJD/studies/35158#
### we'll use this crosswalk to join county to each agency

### import 2012 raw le_agency_crosswalk.rda from icpsr download
load(file=file.path(sharepoint_jr_clearance_le_agency_crosswalk, "le_agency_crosswalk.rda"))

# in 2022 ct change the names and boundaries of their counties
# this will upate the existing crosswalk with a new fips for each ori
# going to upate all the ct agencies with these new fips
ct_county_update <- read_csv(file.path(sp_data_to_plot_path,"ct-county-update-ori.csv")) |> 
  dplyr::select(ori, county_fips = new_county_fips)

### rename and clean dataframe
df_le_agency_crosswalk <- da35158.0001 |> 
  clean_names() |> 
  ### concatenate fips_state_code and fips_county_code 
  ### clean names
  mutate(county_fips = paste0(fips_st,
                              fips_county),
         county = trimws(str_to_title(countyname),
                         which = "both"),
         state_name = trimws(str_to_title(as.character(statename)),
                             which = "both"),
         name = str_to_title(name)) |> 
  dplyr::select(agency_name_lookup = name,
                ori = ori9,
                county,
                county_fips) |> 
  ### NOTE: SOME AGENCIES APPEAR TO HAVE WRONG ORI ID
  ### BIGGEST EXAMPLE IS PHILADELPHIA PD
  ### BASED ON ORI IN THIS CROSSWALK
  ### WE JOIN TO INDIANA COUNTY INSTEAD
  ### OF PHILADELPHIA COUNTY
  ### hard code fix for philly
  ### is this just due to the 2012 file being outdated?
  mutate(county = if_else(ori=="PAPEP0000",
                      "Philadelphia",county),
         county_fips = if_else(ori=="PAPEP0000",
                      "42101",county_fips)) |>
  distinct() |> 
  rows_update(ct_county_update, by = "ori")


### remove original file
rm(da35158.0001)

```

```{r }
################################################################
### pull in county-level population estimates
### we end up just using this file for cleaned up county names
### they appear to be better than the names from UCR/NIBRS
################################################################

### create list of words to exclude from county name -- but only if the exact string matches
### as opposed to a part of another word (i.e., we don't want to remove "and" from "Grand")
words_to_exclude <- c("County","Parish","Census","Municipality","Planning Region", "Area")

words_to_exclude_use <- paste0("\\b", paste(words_to_exclude , collapse="\\b|\\b"), "\\b")

### county-level file
### pulling in population data in case we want to use it as tooltip
us_county_population <- get_acs(
  geography = "county",
  variables = "S0101_C01_001",
  year = 2022,
  geometry = FALSE,
  cb = TRUE,
  resolution = "20m") |>
  clean_names() |>
  mutate(state_geoid = substr(geoid, 1, 2)) |>
  ### extract county and state name and trim any leading/trailing whitespace
  ### need code that accounts for a variety of geographies (AK + LA don't have counties)
  ### may need to continue adding as other non-county examples arise
  mutate(acs_county_name = trimws(str_remove_all(name, 
                                                  regex(words_to_exclude_use, 
                                                        ignore_case = TRUE))),
         acs_county_state_name = trimws(str_extract(name, '\\b[^,]+$')),
         acs_county_name = trimws(str_remove_all(acs_county_name, paste0(", ",acs_county_state_name)))) |>
  dplyr::select(county_fips = geoid,
                acs_county_name,
                acs_county_pop=estimate,
                acs_county_state_name) |>
  distinct()

```

## 1. Clearance by type of violent crime (homicide, rape, robbery, agg assault) by LEA

```{r }
###################################################
### this chunk produces an analytic file
### for the updated section of the site
### with clearance rates by violent crime type
### and % change over last 5 years (2018-2023)
### this syntax is similar to above except that 
### we are additionally grouping by "group" column
###################################################


###################################################
### note: 'group' = type of offense
### create agency- and group-level file for table
### two measures: 2023 unsolved rate 
### and 2015-2023 percent change
#################################################
df_nibrs_violent_unsolved_by_type_year_agency <- df_nibrs_kaplan |> 
  filter(group %in% c("Homicide", "Rape", "Robbery", "Aggravated assault")) |> 
  ### add county from lookup
  left_join(df_le_agency_crosswalk,
            by = "ori") |> 
  dplyr::group_by(year,
                  county,
                  county_fips,
                  agency_name,
                  agency_type,
                  ori,
                  pop_jurisdiction,
                  state_name,
                  state_abbr,
                  number_of_months_missing,
                  number_of_months_reported,
                  group) |> 
  summarize(n_reported = sum(n_reported,
                             na.rm = TRUE),
            n_cleared = sum(n_cleared,
                             na.rm = TRUE),
            n_unsolved = sum(n_unsolved,
                             na.rm = TRUE)) |> 
  ungroup() |> 
  mutate(unsolved_rate = n_unsolved/n_reported,
         n_reported_rate = n_reported/pop_jurisdiction*100000) |> 
  ### recode any unsolved counts or rates less than 0 to 0
  mutate(unsolved_rate = ifelse(unsolved_rate<0 & !is.na(unsolved_rate),
                                     0,unsolved_rate),
         n_unsolved = ifelse(n_unsolved<0 & !is.na(n_unsolved),
                                     0,n_unsolved)) 


### prep df for agency-level table by type of offense
df_nibrs_violent_unsolved_by_agency_type_to_table <- df_nibrs_violent_unsolved_by_type_year_agency |> 
  group_by(ori,
           year,
           group) |> 
  mutate(n_unsolved = unsolved_rate*n_reported,
         n_solved = n_reported - n_unsolved,
         solved_rate = n_solved / n_reported)  |> 
  ungroup() |> 
  dplyr::select(-n_unsolved) |> 
  distinct(ori,
           year,
           group,
           .keep_all = TRUE) |> 
    dplyr::select(year,
                  agency_name,
                  agency_type,
                  ori,
                  county,
                  county_fips,
                  state_name,
                  state_abbr,
                  population = pop_jurisdiction,
                  group,
                  n_reported:solved_rate, 
                  number_of_months_reported)

### first df with everything we need except for calculation of 2019 to 2023 change
df_nibrs_violent_unsolved_by_agency_type_to_table_2023_1 <- df_nibrs_violent_unsolved_by_agency_type_to_table |> 
  filter(year==2023)

### now calculate 2019-2023 change where possible and join to df_nibrs_violent_unsolved_by_agency_to_table_2023_1
### agency must have both years of data for this calculation
df_nibrs_violent_unsolved_by_agency_type_to_table_2023_to_join <- df_nibrs_violent_unsolved_by_agency_type_to_table |> 
  filter(year%in%c(
    2018,
    2023)) |>
  group_by(ori,
           group) |>
  mutate(years_of_data = n_distinct(year)) |>
  filter(years_of_data==2) |>
  ### currently multiplying rate by 100 instead of using scales::percent
  ### may be better way to format within hc
  mutate(solved_rate_2018 = solved_rate[year==2018],
         solved_rate_2023 = solved_rate[year==2023],
         # solved_rate_percent_change_2019_2023 = ((solved_rate_2023 - solved_rate_2019)/solved_rate_2019),
         solved_rate_percent_point_change_2018_2023 = solved_rate_2023 - solved_rate_2018,
         ### convert to percent point (from decimal)
         solved_rate_percent_point_change_2018_2023 = solved_rate_percent_point_change_2018_2023*100) |>
  ungroup() |>
  ### now drop 2018
  ### and keep unsolved rate from 2022
  filter(year==2023) |> 
    dplyr::select(ori,
                  group,
                  solved_rate_percent_point_change_2018_2023) |> 
  mutate(solved_rate_percent_point_change_2018_2023 = ifelse(is.nan(solved_rate_percent_point_change_2018_2023),
                                                       NA,solved_rate_percent_point_change_2018_2023),
         solved_rate_percent_point_change_2018_2023 = ifelse(is.infinite(solved_rate_percent_point_change_2018_2023),
                                                       NA,solved_rate_percent_point_change_2018_2023))


### now join df_nibrs_violent_unsolved_by_agency_to_table_2023 and df_nibrs_violent_unsolved_by_agency_to_table_2023_to_join
df_nibrs_violent_unsolved_by_agency_and_type_to_table_2023 <- left_join(df_nibrs_violent_unsolved_by_agency_type_to_table_2023_1,
                                                               df_nibrs_violent_unsolved_by_agency_type_to_table_2023_to_join,
                                                               by = c("ori",
                                                                      "group")) |> 
  ### pivot long to wide
  ### one row per agency
  pivot_wider(id_cols = c(year,
                  agency_name,
                  agency_type,
                  ori,
                  county,
                  county_fips,
                  state_name,
                  state_abbr,
                  population,
                  number_of_months_reported),
              names_from = group, 
              values_from = c(n_reported,n_reported_rate,n_solved,solved_rate,solved_rate_percent_point_change_2018_2023)) |> 
  ### clean up column names from pivot
  clean_names() |> 
  ### join acs county level file to clean up county name
  ### similar to mapping df used for site
  left_join(us_county_population,
            by = "county_fips") |>
  ### county name from ACS appears cleaner than from lea lookup table
  ### so we'll clean up `county` -- using the acs version when not missing from lookup 
  mutate(county_clean = ifelse(!is.na(acs_county_name),
                               acs_county_name,county)) |> 
    dplyr::select(agency_name,
                  agency_type,
                  ori,
                  county=county_clean,
                  state_name,
                  state_abbr,
                  jurisdiction_population = population,
                  n_reported_aggravated_assault:solved_rate_percent_point_change_2018_2023_robbery, 
                  number_of_months_reported) 



### write out
write_rds(df_nibrs_violent_unsolved_by_agency_and_type_to_table_2023,
          file.path(sp_data_site_updates_path,"website_update_viol_crime_unsolved_by_agency_and_type.rds"))

```

<br><br>

## 2. Resources

### Budget for LEAs over time – 5-10 years (by state: local and State Police Expenditure)

**Note:** All expenditure data has been adjusted to account for income inflation over time using the following formula (e.g. to adjust 1995 data to 2022 numbers):

* 1995 Inflation-adjusted estimate = 1995 estimate * (2022 index value / 1995 index value)

See: https://www.census.gov/topics/income-poverty/income/guidance/current-vs-constant-dollars.html


**Source:** https://www.census.gov/programs-surveys/gov-finances/data/datasets.All.List_1883146942.html#list-tab-List_1883146942

```{r include=FALSE,message=FALSE,warning=FALSE}

### import census survey of local and state gov expenditure
asslgf <- read_rds(file.path(sp_jr_data_library, "census", "asslgf", "census_asslgf_state.rds")) |> 
  ### drop territories
  right_join(states_list,
             by = "state_abbr")


### import inflation index info for calculation of inflation-adjusted expenses
#### from here: https://www.census.gov/topics/income-poverty/income/guidance/current-vs-constant-dollars.html
inflation_index_years <- read_xls(file.path(sp_clearance_rate_data_path,"annual-index-value_annual-percent-change.xls")) |> 
  clean_names() |> 
  dplyr::select(year=income_year,
                inflation_index = index)

### adjust expenditure data to be inflation adjusted
asslgf_clean <- asslgf |> 
  filter(indicator=="Expenditures",
         group=="Police Protection") |> 
  left_join(inflation_index_years,
            by = c("year")) |> 
  ### resource for adjusting for inflation: https://www.census.gov/topics/income-poverty/income/guidance/current-vs-constant-dollars.html
  mutate(n_inflation_adjusted_2022_dollars = n * (163.9 / inflation_index),
         inflation_adjusted_rate_total = n_inflation_adjusted_2022_dollars/pop_total) |> 
  ### keep last 10 years of data
  filter(year>=2012)

### write out
write_rds(asslgf_clean,
          file.path(sp_data_site_updates_path,"website_update_state_level_police_expenditure_inflation_adjusted.rds"))

```

<br>

#### Local and State Police Expenditure as % of Total Budget

```{r include=FALSE,message=FALSE,warning=FALSE}

### now look at police spending as % of total budget
asslgf_percent_clean <- asslgf |> 
  filter(indicator=="Percent of total state expenditure",
         group=="Police Protection") |> 
  ### keep last 10 years of data
  filter(year>=2012)

### write out
write_rds(asslgf_percent_clean,
          file.path(sp_data_site_updates_path,"website_update_state_level_police_expenditure_percent_of_budget.rds"))

```

### Staffing for LEAs over time – 5-10 years


#### UCR: law enforcement employee data 

**Note:** In the following analysis, I'm using the count of sworn officers only, not civilians. Also, Nebraska is removed because they don't appear to have any reported lea employee data in this sample window

**Source:**  https://cde.ucr.cjis.gov/LATEST/webapp/#/pages/downloads

```{r }
### set clean data sharepoint path (for export)
sharepoint_jr_clearance_ucr_employee <- file.path(sp_clearance_rate_data_path,"UCR Data/Employee Data")

### import and clean file
### calculate officers per 10k residents in agency jurisdiction
ucr_employee_data_raw <- readr::read_csv(file.path(sharepoint_jr_clearance_ucr_employee,
                                         "lee_1960_2023.csv")) |> 
  dplyr::select(ori,
                year=data_year,
                state_abbr,
                agency_type_name,
                population,
                officer_ct, 
                civilian_ct) |> 
  filter(year>=2012) |> 
  mutate(officer_pop_rate_per_10k = (officer_ct/population)*10000)

```

<br>


### Workload over time – 5-10 years

*	Officers per capita

*	Officers per violent crime

*	Officers per homicide


```{r }
########################################
### create agency-level file for join
### to employee data
########################################
df_nibrs_violent_unsolved_by_year_agency_offense_type <- df_nibrs_kaplan |> 
  filter(group %in% c("Homicide", "Rape", "Robbery", "Aggravated assault"),
         number_of_months_reported==12,
         year >= 2012) |> 
  dplyr::group_by(year,
                  ori) |> 
  summarize(n_all_violent_reported = sum(n_reported,
                             na.rm = TRUE),
            n_homicide_reported = sum(n_reported[group="Homicide"])) |> 
  ungroup() 

### now calculate officers per violent crime and homicide over time
ucr_employee_data_clean <- left_join(ucr_employee_data_raw,
                                     df_nibrs_violent_unsolved_by_year_agency_offense_type,
                                     by = "ori") |> 
  mutate(officer_per_violent_offense_1k = (officer_ct/n_all_violent_reported)*1000,
         officer_per_homicide_100 = (officer_ct/n_homicide_reported)*100)


### write out
write_rds(ucr_employee_data_clean,
          file.path(sp_data_site_updates_path,"website_update_lea_staffing_workload.rds"))

```

<br>

### Correlation between clearance rates and workload, 2022 only (at LEA level).  Highlight LEAs with higher workloads and lower clearance rates?

**TBD**


<br><br>

## 3. Supplemental homicide report data 

### Descriptives

*	Incidents: clearances, type of clearances, offender and victim counts

*	Offense: location, weapons

*	Victim: relationship with offender, age, gender, race, ethnicity

*	Arrestee: age, gender, race, ethnicity 

*	Highlight characteristics that make cases difficult to solve (e.g., firearms, locations)

<br>

### Disparities (also supplemental homicide report data)

*	In victimization

*	In clearances

*	In case characteristics (e.g., incidents, offense, victim, arrestee)

**Methodological notes:** include the % of homicides represented in the data.  Should this be giant table by LEA or aggregated to county or state levels?

**NOTE:** Any disaggregation by a variable that is correlated with a clearance (e.g., arrestee demographics, weapon used, victim/arresteee relationship, etc.) should be interpreted with caution and included with a disclaimer. 

```{r include=FALSE,message=FALSE,warning=FALSE}
### note: this syntax will only work if you have already set your project path

### set path for use in combination with additional subfolder names
sp_clearance_rate_data_path <- csg_get_project_path("clearance_rates")

### set clean data sharepoint path (for export)
sharepoint_jr_clearance_ucr_clean <- file.path(sp_clearance_rate_data_path,"UCR Data/Clean")

### set path for shr data (through 2022)
sharepoint_jr_clearance_ucr_shr <- file.path(sp_clearance_rate_data_path,"UCR Data/Supplemental Homicide Reports")

```

```{r include=FALSE,message=FALSE,warning=FALSE}
### read in crime data
ucr_srs_state <- read_rds(file.path(sp_data_to_plot_path, "srs_state.rds")) |> 
  filter(group == "Homicide", 
         year == 2022) |> 
  dplyr::select(state_name,
                state_abb = state_abbr,
                incidents_reported)
  
```

```{r include=FALSE,message=FALSE,warning=FALSE}
### read in shr_1976_2022
### downloaded from icprsr: https://www.openicpsr.org/openicpsr/project/100699/version/V14/view?path=/openicpsr/100699/fcr:versions/V14/shr_1976_2022.rds&type=file
shr_1976_2022 <- read_rds(file.path(sharepoint_jr_clearance_ucr_shr, "shr_1976_2022.rds")) 

### clean up shr
### overall
shr_2022_state <- shr_1976_2022 |> 
  filter(year==2022) |> 
  ### calculate clearance rate
  ### code any incident where offender information is populated (e.g. age, sex, or race/ethnicity) as cleared
  mutate(cleared_flag = ifelse(offender_1_age!="unknown" & !is.na(offender_1_age) & 
                                 offender_1_race!="unknown" & !is.na(offender_1_race) &
                                 offender_1_sex!="unknown" & !is.na(offender_1_sex),
                               1,0),
         cleared_flag = ifelse(is.na(cleared_flag),
                               0,cleared_flag)) |> 
  group_by(state_abb,
           year) |> 
  summarize(n_total_incidents = n(),
            n_total_cleared = sum(cleared_flag,na.rm = TRUE),
            clearance_rate = n_total_cleared/n_total_incidents) |> 
  ungroup()

### join shr and ucr data to explore the percent of all ucr reported homicides that appear in SHR
shr_2022_state_ucr <- shr_2022_state |> 
  left_join(ucr_srs_state,
            by = "state_abb") |> 
  mutate(`Percent Coverage (SHR/UCR SRS)` = scales::percent(n_total_incidents/incidents_reported,
                                                            accuracy = 1)) |> 
  dplyr::select(`State` = state_name,
                `Percent Coverage (SHR/UCR SRS)`,
                `SHR Homicides (N)` = n_total_incidents,
                `UCR SRS Homicides (N)` = incidents_reported) |> 
  filter(!is.na(`State`))


# ### export out
# writexl::write_xlsx(shr_2022_state_ucr,
#                     file.path(sharepoint_jr_clearance_av_meeting_viz,
#                               "shr_reported_homicides_ucr_srs_coverage_2022.xlsx"))

### write out
write_rds(shr_2022_state_ucr,
          file.path(sp_data_site_updates_path,
                    "shr_reported_homicides_ucr_srs_coverage_2022.xlsx"))
```


```{r }
### overall
shr_1976_2022_national_overall <- shr_1976_2022 |> 
  ### calculate clearance rate
  ### code any incident where offender information is populated (e.g. age, sex, or race/ethnicity) as cleared
  mutate(cleared_flag = ifelse(offender_1_age!="unknown" & !is.na(offender_1_age) & 
                                 offender_1_race!="unknown" & !is.na(offender_1_race) &
                                 offender_1_sex!="unknown" & !is.na(offender_1_sex),
                               1,0),
         cleared_flag = ifelse(is.na(cleared_flag),
                               0,cleared_flag)) |> 
  group_by(year) |> 
  summarize(n_total_incidents = n(),
            n_total_cleared = sum(cleared_flag,na.rm = TRUE),
            clearance_rate = n_total_cleared/n_total_incidents) |> 
  ungroup()

# ### double check coding for age
# table(shr_1976_2022_by_agency_over_time$offender_1_age[shr_1976_2022_by_agency_over_time$cleared_flag==0], useNA = 'always')
# 
# ### double check coding for race
# table(shr_1976_2022_by_agency_over_time$offender_1_race[shr_1976_2022_by_agency_over_time$cleared_flag==0])
# 
# ### double check coding for sex
# table(shr_1976_2022_by_agency_over_time$offender_1_sex[shr_1976_2022_by_agency_over_time$cleared_flag==0])

### relationship
shr_1976_2022_national_relationship <- shr_1976_2022 |> 
  ### calculate clearance rate
  ### code any incident where offender information is populated (e.g. age, sex, or race/ethnicity) as cleared
  mutate(cleared_flag = ifelse(offender_1_age!="unknown" & !is.na(offender_1_age) & 
                                 offender_1_race!="unknown" & !is.na(offender_1_race) &
                                 offender_1_sex!="unknown" & !is.na(offender_1_sex),
                               1,0),
         cleared_flag = ifelse(is.na(cleared_flag),
                               0,cleared_flag)) |> 
  group_by(year,
           victim_1_relation_to_offender_1) |> 
  summarize(n_total_incidents = n(),
            n_total_cleared = sum(cleared_flag,na.rm = TRUE),
            clearance_rate = n_total_cleared/n_total_incidents) |> 
  ungroup() |> 
  mutate(victim_1_relation_to_offender_1 = str_to_title(victim_1_relation_to_offender_1)) |> 
  dplyr::rename(victim_1_relation_to_offender_1 = victim_1_relation_to_offender_1)

### weapon used
shr_1976_2022_national_weapon <- shr_1976_2022 |> 
  ### calculate clearance rate
  ### code any incident where offender information is populated (e.g. age, sex, or race/ethnicity) as cleared
  mutate(cleared_flag = ifelse(offender_1_age!="unknown" & !is.na(offender_1_age) & 
                                 offender_1_race!="unknown" & !is.na(offender_1_race) &
                                 offender_1_sex!="unknown" & !is.na(offender_1_sex),
                               1,0),
         cleared_flag = ifelse(is.na(cleared_flag),
                               0,cleared_flag)) |> 
  group_by(year,
           offender_1_weapon) |> 
  summarize(n_total_incidents = n(),
            n_total_cleared = sum(cleared_flag,na.rm = TRUE),
            clearance_rate = n_total_cleared/n_total_incidents) |> 
  ungroup() |> 
  mutate(offender_1_weapon = str_to_title(offender_1_weapon)) |> 
  dplyr::rename(offender_weapon_used = offender_1_weapon)

### circumstance of homicide
shr_1976_2022_national_circumstance <- shr_1976_2022 |> 
  ### calculate clearance rate
  ### code any incident where offender information is populated (e.g. age, sex, or race/ethnicity) as cleared
  mutate(cleared_flag = ifelse(offender_1_age!="unknown" & !is.na(offender_1_age) & 
                                 offender_1_race!="unknown" & !is.na(offender_1_race) &
                                 offender_1_sex!="unknown" & !is.na(offender_1_sex),
                               1,0),
         cleared_flag = ifelse(is.na(cleared_flag),
                               0,cleared_flag)) |> 
  group_by(year,
           offender_1_circumstance) |> 
  summarize(n_total_incidents = n(),
            n_total_cleared = sum(cleared_flag,na.rm = TRUE),
            clearance_rate = n_total_cleared/n_total_incidents) |> 
  ungroup() |> 
  mutate(offender_1_circumstance = str_to_title(offender_1_circumstance)) |> 
  dplyr::rename(circumstance_of_homicide = offender_1_circumstance)

### race
shr_1976_2022_national_race <- shr_1976_2022 |> 
  ### calculate clearance rate
  ### code any incident where offender information is populated (e.g. age, sex, or race/ethnicity) as cleared
  mutate(cleared_flag = ifelse(offender_1_age!="unknown" & !is.na(offender_1_age) & 
                                 offender_1_race!="unknown" & !is.na(offender_1_race) &
                                 offender_1_sex!="unknown" & !is.na(offender_1_sex),
                               1,0),
         cleared_flag = ifelse(is.na(cleared_flag),
                               0,cleared_flag)) |> 
    ### clean race/ethnicity and gender
  ### use factor function to assign labels to values
  mutate(victim_1_race_clean = case_when(
    (victim_1_race=="american indian or alaskan native") & victim_1_ethnic_origin!="Hispanic" ~ 1,
    (victim_1_race%in%c("asian","native hawaiian or other pacific islander")) & victim_1_ethnic_origin!="Hispanic" ~ 2,
    (victim_1_race=="black") & victim_1_ethnic_origin!="Hispanic" ~ 3,
    victim_1_ethnic_origin=="Hispanic" ~ 4,
    (victim_1_race=="white") & victim_1_ethnic_origin!="Hispanic" ~ 5,
    victim_1_race=="unknown" & victim_1_ethnic_origin!="Hispanic" ~ 6,
    TRUE ~ 7),
    victim_1_race_clean = factor(victim_1_race_clean,
         levels = c(1,2,3,4,5,6,7),
         labels = c("American Indian",
                    "Asian/Pacific Islander",
                    "Black",
                    "Hispanic",
                    "White",
                    "Unknown",
                    "Missing"))) |> 
  group_by(year,
           victim_1_race_clean) |> 
  summarize(n_total_incidents = n(),
            n_total_cleared = sum(cleared_flag,na.rm = TRUE),
            clearance_rate = n_total_cleared/n_total_incidents) |> 
  ungroup()

### gender
shr_1976_2022_national_gender <- shr_1976_2022 |> 
  ### calculate clearance rate
  ### code any incident where offender information is populated (e.g. age, sex, or race/ethnicity) as cleared
  mutate(cleared_flag = ifelse(offender_1_age!="unknown" & !is.na(offender_1_age) & 
                                 offender_1_race!="unknown" & !is.na(offender_1_race) &
                                 offender_1_sex!="unknown" & !is.na(offender_1_sex),
                               1,0),
         cleared_flag = ifelse(is.na(cleared_flag),
                               0,cleared_flag)) |> 
  mutate(victim_1_sex_clean = case_when(
    victim_1_sex=="female" ~ 1,
    victim_1_sex=="male" ~ 2,
    victim_1_sex=="unknown" ~ 3,
    TRUE ~ as.numeric(NA)),
    victim_1_sex_clean = factor(victim_1_sex_clean,
         levels = c(1,2,3),
         labels = c("Female", "Male", "Unknown"))) |> 
  group_by(year,
           victim_1_sex_clean) |> 
  summarize(n_total_incidents = n(),
            n_total_cleared = sum(cleared_flag,na.rm = TRUE),
            clearance_rate = n_total_cleared/n_total_incidents) |> 
  ungroup()


### race and sex over time
shr_1976_2022_national_race_sex <- shr_1976_2022 |> 
  ### calculate clearance rate
  ### code any incident where offender information is populated (e.g. age, sex, or race/ethnicity) as cleared
  mutate(cleared_flag = ifelse(offender_1_age!="unknown" & !is.na(offender_1_age) & 
                                 offender_1_race!="unknown" & !is.na(offender_1_race) &
                                 offender_1_sex!="unknown" & !is.na(offender_1_sex),
                               1,0),
         cleared_flag = ifelse(is.na(cleared_flag),
                               0,cleared_flag)) |> 
    ### clean race/ethnicity and gender
  ### use factor function to assign labels to values
  mutate(victim_1_race_clean = case_when(
    (victim_1_race=="american indian or alaskan native") & victim_1_ethnic_origin!="Hispanic" ~ 1,
    (victim_1_race%in%c("asian","native hawaiian or other pacific islander")) & victim_1_ethnic_origin!="Hispanic" ~ 2,
    (victim_1_race=="black") & victim_1_ethnic_origin!="Hispanic" ~ 3,
    victim_1_ethnic_origin=="Hispanic" ~ 4,
    (victim_1_race=="white") & victim_1_ethnic_origin!="Hispanic" ~ 5,
    victim_1_race=="unknown" & victim_1_ethnic_origin!="Hispanic" ~ 6,
    TRUE ~ 7),
    victim_1_race_clean = factor(victim_1_race_clean,
         levels = c(1,2,3,4,5,6,7),
         labels = c("American Indian",
                    "Asian/Pacific Islander",
                    "Black",
                    "Hispanic",
                    "White",
                    "Unknown",
                    "Missing"))) |> 
  mutate(victim_1_sex_clean = case_when(
    victim_1_sex=="female" ~ 1,
    victim_1_sex=="male" ~ 2,
    victim_1_sex=="unknown" ~ 3,
    TRUE ~ as.numeric(NA)),
    victim_1_sex_clean = factor(victim_1_sex_clean,
         levels = c(1,2,3),
         labels = c("Female", "Male", "Unknown"))) |> 
  group_by(year,
           victim_1_race_clean,
           victim_1_sex_clean) |> 
  summarize(n_total_incidents = n(),
            n_total_cleared = sum(cleared_flag,na.rm = TRUE),
            clearance_rate = n_total_cleared/n_total_incidents) |> 
  ungroup()


### age
shr_1976_2022_national_age <- shr_1976_2022 |> 
  ### calculate clearance rate
  ### code any incident where offender information is populated (e.g. age, sex, or race/ethnicity) as cleared
  mutate(cleared_flag = ifelse(offender_1_age!="unknown" & !is.na(offender_1_age) & 
                                 offender_1_race!="unknown" & !is.na(offender_1_race) &
                                 offender_1_sex!="unknown" & !is.na(offender_1_sex),
                               1,0),
         cleared_flag = ifelse(is.na(cleared_flag),
                               0,cleared_flag)) |> 
  ### create age bins
   mutate(victim_1_age_bin = case_when(
    victim_1_age<25 ~ 1,
    victim_1_age>=25 & victim_1_age<=34 ~ 2,
    victim_1_age>34 & victim_1_age<=45 ~ 3,
    victim_1_age>45 ~ 4,
    TRUE ~ as.numeric(NA)),
    victim_1_age_bin = factor(victim_1_age_bin,
         levels = c(1,2,3,4),
         labels = c("18 to 24", "25 to 34", "35 to 45", "46+"))) |> 
  group_by(year,
           victim_1_age_bin) |> 
  summarize(n_total_incidents = n(),
            n_total_cleared = sum(cleared_flag,na.rm = TRUE),
            clearance_rate = n_total_cleared/n_total_incidents) |> 
  ungroup()


# ### arrestee race
# shr_1976_2022_national_arrestee_race <- shr_1976_2022 |> 
#   ### calculate clearance rate
#   ### code any incident where offender information is populated (e.g. age, sex, or race/ethnicity) as cleared
#   mutate(cleared_flag = ifelse(offender_1_age!="unknown" & !is.na(offender_1_age) & 
#                                  offender_1_race!="unknown" & !is.na(offender_1_race) &
#                                  offender_1_sex!="unknown" & !is.na(offender_1_sex),
#                                1,0),
#          cleared_flag = ifelse(is.na(cleared_flag),
#                                0,cleared_flag)) |> 
#   ### clean race/ethnicity and gender
#   ### use factor function to assign labels to values
#   mutate(offender_1_race_clean = case_when(
#     (offender_1_race=="american indian or alaskan native") & offender_1_ethnic_origin!="Hispanic" ~ 1,
#     (offender_1_race%in%c("asian","native hawaiian or other pacific islander")) & offender_1_ethnic_origin!="Hispanic" ~ 2,
#     (offender_1_race=="black") & offender_1_ethnic_origin!="Hispanic" ~ 3,
#     offender_1_ethnic_origin=="Hispanic" ~ 4,
#     (offender_1_race=="white") & offender_1_ethnic_origin!="Hispanic" ~ 5,
#     offender_1_race=="unknown" & offender_1_ethnic_origin!="Hispanic" ~ 6,
#     TRUE ~ 7),
#     offender_1_race_clean = factor(offender_1_race_clean,
#          levels = c(1,2,3,4,5,6,7),
#          labels = c("American Indian",
#                     "Asian/Pacific Islander",
#                     "Black",
#                     "Hispanic",
#                     "White",
#                     "Unknown",
#                     "Missing"))) |> 
#   group_by(year,
#            offender_1_race_clean) |> 
#   summarize(n_total_incidents = n(),
#             n_total_cleared = sum(cleared_flag,na.rm = TRUE),
#             clearance_rate = n_total_cleared/n_total_incidents) |> 
#   ungroup()
# 
# ### arrestee gender
# shr_1976_2022_national_arrestee_gender <- shr_1976_2022 |> 
#   ### calculate clearance rate
#   ### code any incident where offender information is populated (e.g. age, sex, or race/ethnicity) as cleared
#   mutate(cleared_flag = ifelse(offender_1_age!="unknown" & !is.na(offender_1_age) & 
#                                  offender_1_race!="unknown" & !is.na(offender_1_race) &
#                                  offender_1_sex!="unknown" & !is.na(offender_1_sex),
#                                1,0),
#          cleared_flag = ifelse(is.na(cleared_flag),
#                                0,cleared_flag)) |> 
#   mutate(offender_1_sex_clean = case_when(
#     offender_1_sex=="female" ~ 1,
#     offender_1_sex=="male" ~ 2,
#     offender_1_sex=="unknown" ~ 3,
#     TRUE ~ as.numeric(NA)),
#     offender_1_sex_clean = factor(offender_1_sex_clean,
#          levels = c(1,2,3),
#          labels = c("Female", "Male", "Unknown"))) |> 
#   group_by(year,
#            offender_1_sex_clean) |> 
#   summarize(n_total_incidents = n(),
#             n_total_cleared = sum(cleared_flag,na.rm = TRUE),
#             clearance_rate = n_total_cleared/n_total_incidents) |> 
#   ungroup()

```

```{r }
### now write all dfs out for visualization

### write out
write_rds(shr_1976_2022_national_overall,
          file.path(sp_data_site_updates_path,"shr_1976_2022_national_overall_clear.rds"))

### write out
write_rds(shr_1976_2022_national_relationship,
          file.path(sp_data_site_updates_path,"shr_1976_2022_national_victim_offender_relationship.rds"))

### write out
write_rds(shr_1976_2022_national_weapon,
          file.path(sp_data_site_updates_path,"shr_1976_2022_national_weapon.rds"))

### write out
write_rds(shr_1976_2022_national_circumstance,
          file.path(sp_data_site_updates_path,"shr_1976_2022_national_incident_circumstance.rds"))

### write out
write_rds(shr_1976_2022_national_race,
          file.path(sp_data_site_updates_path,"shr_1976_2022_national_victim_race_eth.rds"))

### write out
write_rds(shr_1976_2022_national_gender,
          file.path(sp_data_site_updates_path,"shr_1976_2022_national_victim_gender.rds"))

### write out
write_rds(shr_1976_2022_national_race_sex,
          file.path(sp_data_site_updates_path,"shr_1976_2022_national_victim_race_gender.rds"))

### write out
write_rds(shr_1976_2022_national_age,
          file.path(sp_data_site_updates_path,"shr_1976_2022_national_victim_age.rds"))

# ### write out
# write_rds(shr_1976_2022_national_overall,
#           file.path(sp_data_site_updates_path,"shr_1976_2022_national_arrestee_race_eth.rds"))
# 
# ### write out
# write_rds(shr_1976_2022_national_overall,
#           file.path(sp_data_site_updates_path,"shr_1976_2022_national_arrestee_gender.rds"))
# 
# ### write out
# write_rds(shr_1976_2022_national_overall,
#           file.path(sp_data_site_updates_path,"shr_1976_2022_national_arrestee_age.rds"))


```

<br><br>
